### 2018-04-17 MSLT/GBDCompare.Rmd
Separate compare plots to another doc

---
title: "GBDCompare"
author: "Belen Zapata Diomedi"
date: "17 April 2018"
output: pdf_document
---

The following graphs compare rates per one by age (5-year groups) and sex for England and Greater London.

```{r, echo = FALSE}
###TO DO: cubic splines for input rates

#Global options

#in_data1: GBDEngland
#in_data2: GBDGL
require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
source("code/functions.R")

GBDdata <- read.csv("data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
GBDdata <- mutate_all(GBDdata, funs(tolower))

GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97

GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )

GBDdata$val <- as.numeric(as.character(GBDdata$val))

GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)

GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
GBDEngland$one_rate <- GBDEngland$val/100000
GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
GBDGL$one_rate <- GBDGL$val/100000

p_sex <- c("male", "female")
p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancer", "breast cancer", "tracheal, bronchus, and lung cancer")
p_measure <- c("deaths", "ylds (years lived with disability)", "prevalence", "incidence")

for (sex in p_sex){
  for (cause in p_cause) {
    for (measure in p_measure){

# Exclude breast_cancer for Males
      if (sex == "male" && cause == "breast cancer" || cause == "all causes" && measure == "prevalence" || cause == "all causes" && measure == "incidence"){
        p1 <- NULL

      }
      else {
      p1 <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure)

      }
    }
  }
}

```