### 2018-07-09 doc/accra/process_flow.rmd
Add rmd doc for accra process flow

---
title: "Accra - process flow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario Creation - AA

Read leandro’s created processed raw data (which already has domain specific duration and days) as:

rd <- read.csv("data/synth_pop_data/accra/travel_survey/synthetic_population_with_trips.csv")

Leandro has already removed >70 population from it
This has people without trips (trip_mode 99 and trip_id 0) as well

### Walk trips to Baseline

* Add walk trips out of bus trips to baseline

### Create scenarios

* Scenario 1: Shift modes - 50% of all walking walking trips to Private Car

* Scenario 1: Shift modes - 50% of all car trips less than 7km to cycle

* Scenario 3: Shift and create new trips - 50% of all trips longer than 10km to Bus (as each bus trip might create a new short walking trip)

* Save scenarios as: data/scenarios/accra/baseline_and_three_scenarios.csv
(contains mode and duration of each trips for baseline and all three scenarios)

## Physical Activity – AA

Calculate total PA per individual
1.	Multiply trips with AT by 6 (5 * 1.2) to get weekly trips. 
2.	Sum them for each individual to get total PA

## Air Pollution – RG
Fill in the info

## Health – RG

Identify PA and AP related diseases and then calculate their RR.

Multiply PA’s and AP’s RR to get combined effect

Normalize them – see a message from James below:

*When adjusting AP for PA I think we should renormalize the values if we are using non-linear RRs. The epid studies didn’t allow for PA levels so they will be coming up with lower equivalent levels than we are. Thus I would suggest that for an average individual (how to define that is a challenge) we estimate the ratio of our modelled inhaled exposure to background and apply this to everyone. Does that make sense?*
*The same would apply for diurnal patterns etc.*

### Calculate disease specific burden 
* Calculate PIFs – CRA for baseline and Scenario
* Calculate disease burden (using GBD’s data) for deaths and YLLs