### 2019-05-13 diagnostic/dr_curves.Rmd
dose--response curves added and appended below parameter_table_all

---
title: "Dose--response curves"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(ithimr)
```

```{r pa, include=FALSE}

global_path <- file.path(find.package('ithimr',lib.loc=.libPaths()), 'extdata/global/')
global_path <- paste0(global_path, "/")

DISEASE_INVENTORY <- read.csv(paste0(global_path,"dose_response/disease_outcomes_lookup.csv"))

list_of_files <- list.files(path = paste0(global_path,"dose_response/drpa/extdata/"), recursive = TRUE, pattern = "\\.csv$", full.names = TRUE)
for (i in 1:length(list_of_files)){
  assign(stringr::str_sub(basename(list_of_files[[i]]), end = -5),
         readr::read_csv(list_of_files[[i]],col_types = cols()),
         pos = 1)
}

########## dose response

doses_vector <- as.numeric(0:50)
cols <- rainbow(sum(DISEASE_INVENTORY$physical_activity == 1))
cause_indices <- which(DISEASE_INVENTORY$physical_activity == 1)

############# median dose-response curves
# PA_DOSE_RESPONSE_QUANTILE <<- 0.5
# #pdf('results/dose_response/PA_dose_response.pdf',width=11,height=11);par(mar=c(5,5,2,1))
# for ( j in 1:length(cause_indices)){
#   cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
#   # get name of disease
#   print(cause)
#   return_vector <- PA_dose_response(cause = cause,dose = doses_vector)
#   
#   if(j==1){
#     plot(doses_vector,return_vector$rr,type='l',ylim=0:1,main='',ylab='Relative risk',xlab='Dose',
#          frame=F,col=cols[j],xlim=range(doses_vector),cex.axis=1.5,cex.lab=1.5,cex.main=1.5,lwd=2) 
#   }else{
#     lines(doses_vector,return_vector$rr,col=cols[j],lwd=2)
#   }
# }
# legend(legend=DISEASE_INVENTORY$pa_acronym[DISEASE_INVENTORY$physical_activity==1],col=cols,bty='n',x=0,y=0.5,lwd=2)
# #dev.off()
# 
# ############# sample dose-response curves
# nsamples <- 10
# parameters <- ithim_setup_parameters(NSAMPLES=nsamples,
#                                      PA_DOSE_RESPONSE_QUANTILE=T)
# #pdf('results/dose_response/PA_dose_response_sample.pdf',width=11,height=11);
# par(mar=c(5,5,2,1),mfrow=c(3,3))
# for ( j in 1:length(cause_indices)){
#   cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
#   # get name of disease
#   print(cause)
#   for(seed in 1:nsamples){
#     for(i in 1:length(parameters))
#       assign(names(parameters)[i],parameters[[i]][[seed]],pos=1)
#     return_vector <- PA_dose_response(cause = cause,dose = doses_vector)
#     
#     if(seed==1){
#       plot(doses_vector,return_vector$rr,type='l',ylim=0:1,main=cause,ylab='Relative risk',xlab='Dose',
#            frame=F,col=cols[j],xlim=range(doses_vector),cex.axis=1.5,cex.lab=1.5,cex.main=1.5,lwd=1) 
#     }else{
#       lines(doses_vector,return_vector$rr,col=cols[j],lwd=1)
#     }
#   }
# }
# #dev.off()
# 
# ################ explanatory graph
# library(diagram)
# j <- 3
# ylow <- 0.6
# doses_clean <- doses_vector
# {x11(); par(mfrow=c(2,2),mar=c(2,5,1,1))
#   pa_dn <- as.character(DISEASE_INVENTORY$pa_acronym[j])
#   pa_n <- as.character(DISEASE_INVENTORY$acronym[j])
#   outcome_type <- ifelse(pa_dn%in%c('lung_cancer','stroke'), 'incidence' , 'mortality')
#   dose <- doses_clean
#   if(pa_dn %in% c('total_cancer','coronary_heart_disease')) dose[dose>35] <- 35
#   cause<-pa_dn
#   fname <- paste(cause, outcome_type, sep = "_")
#   lookup_table <- get(paste0(fname))
#   lookup_df <- as.data.frame(lookup_table)
#   rr <- approx(x=lookup_df$dose,y=lookup_df$RR,xout=dose,yleft=1,yright=min(lookup_df$RR))$y
#   lb <- approx(x = lookup_df$dose,y = lookup_df$lb,xout = dose,yleft = 1,yright = min(lookup_df$lb))$y
#   ub <-approx(x = lookup_df$dose,y = lookup_df$ub,xout = dose,yleft = 1,yright = min(lookup_df$ub))$y
#   plot(doses_clean,rr,lwd=2,type='l',col=rgb(0.6,0.2,0.2),
#        xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='')
#   lines(doses_clean,lb,lwd=2,type='l',lty=2,col=rgb(0.6,0.2,0.2,1))
#   lines(doses_clean,ub,lwd=2,type='l',lty=2,col=rgb(0.6,0.2,0.2,1))
#   text('Mean',x=40,y=0.8,col=rgb(0.6,0.2,0.2),cex=1.5)
#   text('Lower bound',x=35,y=0.66,col=rgb(0.6,0.2,0.2),cex=1.5)
#   text('Upper bound',x=35,y=0.93,col=rgb(0.6,0.2,0.2),cex=1.5)
#   par(mar=c(2,2,1,1))
#   for(i in 1:500){
#     x<-runif(1)
#     rr0 <- truncnorm::qtruncnorm(p=x, mean=rr, sd=(rr-lb)/1.96,a=0)#a=lb,b=ub)#
#     if(i==1){plot(doses_clean,rr0,type='l',col=rgb(0.6,0.2,0.2,0.1),
#                   xlim=range(doses_clean),ylim=c(ylow,1),ylab='',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='') 
#       text('Samples',x=35,y=0.97,col=rgb(0.6,0.2,0.2),cex=1.5)
#     }else lines(doses_clean,rr0,col=rgb(0.6,0.2,0.2,0.1))
#   }
#   quant <- 0.5
#   par(mar=c(5,5,1,1))
#   rr0 <- truncnorm::qtruncnorm(p=quant, mean=rr, sd=(rr-lb)/1.96,a=0)
#   plot(doses_clean,rr0,lwd=2,type='l',col='grey',
#        xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='Dose')
#   doses <- c(5,25,45)
#   text('Median',x=10,y=0.97,col='grey',cex=1.5)
#   abline(h=ylow,lwd=2,lty=2,col='grey')
#   points(doses,rep(ylow,length(doses)),col=rgb(0.6,0.2,0.2),pch=18,cex=1.5)
#   points(doses,rr0[which(doses_clean%in%doses)],col=rgb(0.6,0.2,0.2),pch=15,cex=1.5)
#   for(dose in doses) arrows(x0=dose,y0=ylow,y1=rr0[which(doses_clean==dose)]-0.02,col=rgb(0.6,0.2,0.2),length=0.1,lty=1,lwd=2)
#   par(mar=c(5,2,1,1))
#   cols <- c('hotpink','navyblue','darkorange2')
#   quants <- c(0.1,0.6,0.9)
#   for(i in 1:length(quants)){
#     quant <- quants[i]
#     rr0 <- truncnorm::qtruncnorm(p=quant, mean=rr, sd=(rr-lb)/1.96,a=0)
#     if(i==1) {plot(doses_clean,rr0,lwd=2,type='l',col=cols[i],
#                    xlim=range(doses_clean),ylim=c(ylow,1),ylab='',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='Dose')
#       for(dose in doses) arrows(x0=dose,y0=ylow,y1=rr0[which(doses_clean==dose)]-0.02,col=cols[i],length=0.1,lty=1,lwd=2)
#     }  else{
#       lines(doses_clean,rr0,lwd=2,type='l',col=cols[i])
#       for(dose in doses) 
#         curvedarrow(from=c(dose,ylow),to=c(dose,rr0[which(doses_clean==dose)]-0.01),lty=1,lwd=2,lcol=cols[i],
#                     arr.col=cols[i],arr.pos=0.65+0.05*i,curve=-70+30*i,arr.type='curved',arr.adj=0,endhead=T)
#     }
#     abline(h=ylow,lwd=2,lty=2,col='grey')
#     points(doses,rr0[which(doses_clean%in%doses)],col=cols[i],pch=15,cex=1.5)
#     
#     points(doses,rep(ylow,length(doses)),col=rgb(0.6,0.2,0.2),pch=18,cex=1.5)
#   }
# }

#### plot to html

ylow <- 0.2
doses_clean <- as.numeric(0:60)
```

## PA dose--response curves (4 sd)

```{r plot pa, include=T}
par(mar=c(5,5,2,1),mfrow=c(3,3))
for ( j in 1:length(cause_indices)){
  cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
  # get name of disease
  # decide whether to use "all" or "mortality"
  outcome_type <- ifelse(cause%in%c('lung_cancer','breast_cancer','endometrial_cancer','colon_cancer'), 'all' , 'mortality')

  dose <- doses_clean
  # apply disease-specific thresholds
  if(cause %in% c('total_cancer','coronary_heart_disease','breast_cancer','endometrial_cancer','colon_cancer')) dose[dose>35] <- 35
  else if(cause == 'lung_cancer') dose[dose>10] <- 10
  else if(cause == 'stroke') dose[dose>32] <- 32
  else if(cause == 'all_cause') dose[dose>16.08] <- 16.08
  fname <- paste(cause, outcome_type, sep = "_")
  lookup_table <- get(paste0(fname))
  lookup_df <- as.data.frame(lookup_table)
  rr <- approx(x=lookup_df$dose,y=lookup_df$RR,xout=dose,yleft=1,yright=min(lookup_df$RR))$y
  lb <- approx(x = lookup_df$dose,y = lookup_df$lb,xout = dose,yleft = 1,yright = min(lookup_df$lb))$y
  ub <-approx(x = lookup_df$dose,y = lookup_df$ub,xout = dose,yleft = 1,yright = min(lookup_df$ub))$y
  plot(doses_clean,rr,lwd=2,type='l',col='navyblue',main=cause,xlab='MMET per week',
       xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5)
  lines(doses_clean,lb,lwd=2,type='l',lty=1,col='grey')
  lines(doses_clean,ub,lwd=2,type='l',lty=1,col='grey')
  #text('Mean',x=40,y=0.8,col=rgb(0.6,0.2,0.2),cex=1.5)
  #text('Lower bound',x=35,y=0.66,col=rgb(0.6,0.2,0.2),cex=1.5)
  #text('Upper bound',x=35,y=0.93,col=rgb(0.6,0.2,0.2),cex=1.5)

}

```

```{r get ap, include=F}
nsamples <- 100
DR_AP <- read.csv(paste0(global_path,"dose_response/drap/dose_response.csv"))
parameters <- ithim_setup_parameters(NSAMPLES=nsamples,
                                     AP_DOSE_RESPONSE_QUANTILE=T)
for(i in 1:length(parameters))
  assign(names(parameters)[i],parameters[[i]],pos=1)
causes <- unique(DR_AP$cause_code)

############## plot parameter samples
dr_ap_list <- list()
for ( j in c(1:nrow(DISEASE_INVENTORY))[DISEASE_INVENTORY$air_pollution == 1]){ 
  cause <- as.character(DISEASE_INVENTORY$ap_acronym[j])
  dr_ap <- subset(DR_AP,cause_code==cause)
  ages <- unique(dr_ap$age_code)
  dr_ap_list[[cause]] <- list()
  #print(cause)
  #print(ages)
  for(age in ages){
    dr_ap_list[[cause]][[age]] <- list()
    dr_ap_age <- subset(dr_ap,age_code==age)
    alpha <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$alpha)
    beta <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$beta)
    gamma <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$gamma)
    tmrel <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$tmrel)
    dr_ap_list[[cause]][[age]]$alpha <- alpha
    dr_ap_list[[cause]][[age]]$beta <- beta
    dr_ap_list[[cause]][[age]]$gamma <- gamma
    dr_ap_list[[cause]][[age]]$tmrel <- tmrel
    #pdf(paste0('results/dose_response/AP/',cause,age,'.pdf'))
    #x11(width=8); 
    #par(mfrow=c(2,3),mar=c(4,4,1,1))
    #plot(log(dr_ap_age$alpha),log(dr_ap_age$beta),main=paste0(cause,age),col='grey'); points(log(alpha),log(beta),col='hotpink',pch=18)
    #plot(log(dr_ap_age$alpha),(log(dr_ap_age$gamma)),main=paste0(cause,age),col='grey'); points(log(alpha),log(gamma),col='hotpink',pch=18)
    #plot(log(dr_ap_age$alpha),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(alpha),log(tmrel),col='hotpink',pch=18)
    #plot(log(dr_ap_age$beta),log(dr_ap_age$gamma),main=paste0(cause,age),col='grey'); points(log(beta),log(gamma),col='hotpink',pch=18)
    #plot(log(dr_ap_age$beta),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(beta),log(tmrel),col='hotpink',pch=18)
    #plot(log(dr_ap_age$gamma),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(gamma),log(tmrel),col='hotpink',pch=18)
    #dev.off()
  }
}

```

## AP dose&ndash;response curves (95% intervals)

```{r plot ap}
pm <- 1:300
for ( j in 1:nrow(DISEASE_INVENTORY)) if (DISEASE_INVENTORY$air_pollution[j] == 1){ 
  cause <- as.character(DISEASE_INVENTORY$ap_acronym[j])
  dr_ap <- subset(DR_AP,cause_code==cause)
  ages <- unique(dr_ap$age_code)
  #print(cause)
  #print(ages)
  if(length(ages)>1){ 
    #pdf(paste0('results/dose_response/AP/',cause,'DR-sim.pdf'),width=6,height=8); 
    par(mfrow=c(1,1))
    layout.matrix <- matrix(1:length(ages),nrow=3,ncol=5,byrow=T)
    graphics::layout(mat=layout.matrix,heights=c(4,3,4),widths=c(3.8,3,3,3,3))
  }else{
    #pdf(paste0('results/dose_response/AP/',cause,'DR-sim.pdf'))
    par(mfrow=c(1,1))
  }
  for(age in ages){
    par(mar=c(ifelse(age>70,5,1),ifelse(age%in%c(99,25,50,75),5,1),ifelse(age%in%c(25:45,99),5,1),1))
    dr_ap_age <- subset(dr_ap,age_code==age)
    alpha <- dr_ap_list[[cause]][[age]]$alpha
    beta <- dr_ap_list[[cause]][[age]]$beta
    gamma <- dr_ap_list[[cause]][[age]]$gamma
    tmrel <- dr_ap_list[[cause]][[age]]$tmrel
    pmcurve <- matrix(0,nrow=nsamples,ncol=length(pm))
    for(i in 1:nsamples){#length(alpha)){
      pmcurve[i,] <- ap_dose_response_curve(pm,alpha[i],beta[i],gamma[i],tmrel[i])
      pmcurve[i,is.na(pmcurve[i,])] <- 1
    }
    pmcurvemean <- colMeans(pmcurve)
    pmcurveupper <- apply(pmcurve,2,quantile,0.975)
    pmcurvelower <- apply(pmcurve,2,quantile,0.025)
    plot(pm,pmcurvemean,typ='l',col='navyblue',ylim=c(1,4),main='',cex.lab=1.5,lwd=2,frame=F,xlab=ifelse(age>70,'PM dose',''),ylab=ifelse(age%in%c(99,25,50,75),'RR',''),xaxt='n',yaxt='n',cex.main=1.5)
    axis(1,labels=ifelse(age>70,T,F),cex.axis=1.5)
    axis(2,labels=ifelse(age%in%c(99,25,50,75),T,F),cex.axis=1.5)
    if(length(ages)>1) mtext(side=3,text=paste0('Age ',age),line=0)
    if(age%in%c(35,99)) mtext(side=3,text=as.character(DISEASE_INVENTORY$GBD_name)[j],line=1)
    lines(pm,pmcurveupper,col='grey',lwd=2,cex.axis=1.5)
    lines(pm,pmcurvelower,col='grey',lwd=2,cex.axis=1.5)
  }
  #dev.off()

}

```