### 2019-08-05 code/trips_analysis/trip_distance.Rmd
Markdown document for trip distance analysis

---
title: "Distance distribution by modes"
output: html_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
options(warn = -1)

```

```{r loadLibraries, echo = F, message = F}
suppressWarnings({

  library(tidyverse)
  library(ggthemes)
  library(plotly)
  library(knitr)

})

```

```{r loadData, message = F, echo = F, warning = F}

  io <- read_rds("../../results/multi_city/io.rds")
```

```{r genPlots, message = T, echo = F, warning = F}

suppressWarnings({
  require(knitr)

  l <- htmltools::tagList()
  tl <- list()
  index <- 1

  for (cities in c('accra','sao_paulo','delhi')){

  rd <- io[[cities]]$trip_scen_sets %>% filter(scenario == "Baseline")

  rd <- rd %>% filter(trip_mode %in% c("bus", "walking", "car", "motorcycle", "bicycle"))

  # ggplot(rd) +
  #   aes(x = trip_distance) +
  #   geom_density(adjust = 1L) +
  #   scale_fill_hue() +
  #   theme_tufte() +
  #   facet_wrap(vars(trip_mode), scales = "free") +
  #   labs(title = "Distribution of trips by distance - Accra")

  p <- ggplot(rd) +
  aes(x = trip_mode, y = trip_distance) +
  geom_boxplot(adjust = 1L, scale = "area", fill = "#39486b") +
  geom_hline(yintercept = c(2, 6), linetype="dashed", color = "red", size=0.1) +
  labs(title = paste("Distribution of trips by distance - ", cities), x = "Mode", y = "Distance (km)")

  if (cities == 'delhi'){
    p <- p + coord_cartesian(ylim = c(0, 100)) 
  }

  tl[[index]] <-  filter(p$data, scenario == 'Baseline') %>% select(trip_mode, trip_distance) %>% rename(var = trip_distance) %>% group_by(trip_mode) %>%  summarise(mean = mean(var), median =median(var), Q1 = quantile (var, probs=0.25), Q2 = quantile (var, probs=0.50), Q3 = quantile(var, probs=0.75)) %>% mutate_if(is.numeric, round, digits = 1)

  pp <- plotly_build(p)

  pp$x$data <- lapply(pp$x$data, FUN = function(x){
    x$marker = list(opacity = 0)
    return(x)
  })

  l[[index]] <- as_widget(ggplotly(p))
  index <- index + 1

  }

})

```

## Plots

```{r printPlots, echo = F, message = F}

suppressWarnings({
 l
})
```

## Tables

```{r printTables, echo = F, results = 'asis'}

suppressWarnings({
  require(knitr)
  local_cities <- c('Accra','Sao Paulo','Delhi')
   for (i in 1:3){
      print(kable(tl[[i]], caption = paste("Distribution of trips by distance - ", local_cities[i])))
      cat('\n')
   }
})
```