### 2019-11-04 project_mgmt/data_source_summary.Rmd
summarise data sources

---
title: "Data source summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a summary of data sources as saved in project_mgmt/case_cities_data.xlsx

```{r start, echo=FALSE}
library(xlsx)
#setwd('project_mgmt/')
x <- xlsx::read.xlsx2('case_cities_data.xlsx',sheetIndex=1,colClasses = 'character',stringsAsFactors=FALSE)

start_index <- which(x[,1]=='Country') + 1
stop_index <- which(x[,1]=='Extras') - 1

countries <- x[start_index:stop_index,1]
cities <- x[start_index:stop_index,2]

```

## Travel

```{r travel, echo=FALSE}

travel_section <- which(colnames(x)=='Background.travel.data')
travel_year <- which(x[1,travel_section:ncol(x)]=='Year of data collection')[1] + travel_section - 1
travel_source <- which(x[1,travel_section:ncol(x)]=='Source of data')[1] + travel_section - 1

for(i in start_index:stop_index){
  cat('Travel data for ')
  cat(paste0(cities[i-start_index+1],' (',countries[i-start_index+1],') '))
  cat('for year ')
  cat(x[i,travel_year])
  cat(' were obtained from ')
  cat(x[i,travel_source])
  cat('.\n')
}
```