### 2020-03-19 summary_tables_outcomes.Rmd
Comment out YLLs

---
title: "Summary Tables"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
library(knitr)
library(pracma)
library(summarytools)
library(dplyr)
library(purrr)
library(DT)
library(ggplot2)
library(tidyr)
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')

```

```{r loadLibraries, echo = F, message = F}

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
# much better results. Your mileage may vary

```

```{r load_objects = "asis"}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota','cape_town', 'vizag')

#cities <- c('accra','sao_paulo','buenos_aires')

round_to <- 1

# Set plot_mmets to F
plot_mmets <- F

sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

```

```{r preprocessing}

collapse_ages <- function(data,ages=c('15-49','50-69'),age_lab='age_cat'){
  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
  genders <- unique(data$sex)
  if(ncol(data)>3) {
    reformatted <- do.call(rbind,lapply(1:length(ages),
                                        function(x) t(sapply(genders,
                                                             function(y) 
                                                               if(ncol(data)>3) colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                        ))
    ))
    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
  }else{
    reformatted <- do.call(c,lapply(1:length(ages),
                                    function(x) sapply(genders,
                                                       function(y) 
                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                    )
    ))
    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
  }
}

scen_prop <- io$scen_prop
io <- io[-2]
for(i in 1:length(io)) {
  io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$deaths <- io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}
for(city in cities)
  for(type in c('hb','pathway_hb'))
    for(out in c('deaths','ylls'))
      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]])
for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')
pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population)))
pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj')
ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
scen_names <- rownames(scen_prop)

```

## Who Hit Whom matrix

Who hit whom matrix for all scenarios for each case study

### Who hit whom matrix {#whw}

```{r whw}

scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
scen_lt <- data.frame("scen_name" = names(io$accra$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)

all_whw <- list()

for(city in cities){
  # if (!city %in% c('buenos_aires', 'mexico_city'))
  {
    # city <- 'buenos_aires'
    for (cs in names(io$accra$outcomes$whw)){
      print(city)

      if (is.null(io[[city]]$outcomes$whw[[cs]]$whw)){
        if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
          td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
          td2$mode <- 'NOV'

          td2 <- td2 %>% dplyr::select(mode, names(.))

          td3 <- td2
        }

      }else{
        td1 <- (io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
        td3 <- td1
        if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
          td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
          td2$mode <- 'NOV'

          td3 <- plyr::rbind.fill(td1, td2)
        }
      }
      td3 <- td3 %>% mutate(rowSum = rowSums(.[2:ncol(td3)], na.rm = T))
      td3 <- td3 %>% janitor::adorn_totals("row")
      td3[, 2:ncol(td3)] <- round(td3[, 2:ncol(td3)], 2)

      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()

      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', var, ' matrix for ', city)))
      cat("\n")

      td <- td3
      names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")

      all_whw[[var]][[city]] <- td
    }

  }
}

```

### Who hit whom matrix for all case cities {#whw_all}

```{r whw_for_all_cs, fig.width=7, fig.height=4, echo=FALSE}

st <- list()

file_list <- list()

td <- NULL

for (cs in names(all_whw)){
  # cs <- names(all_whw)[1]

  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = "mode") %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))

  td[is.na(td)] <- 0

  unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unknown')

  unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unspecified')

  other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'other')

  pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'pickup')

  bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'bus')

  td <- td %>% filter(!mode %in% c('bus', 'bus_driver', 'unknown', 'unk', 'unspecified', 'listed_na', '?', 
                                   'other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle',
                                   'picku', 'pickup/light goods'))

  td <- plyr::rbind.fill(td, bus, unk, unsp, other, pu)

  td <- (td[!duplicated(td), ])

  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total'))

  td[is.na(td)] <- 0

  td <- td %>% mutate_if(is.numeric, round, 2)

  backup_td <- td

  readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))

  colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))

  whw_lng <- reshape2::melt(td)

  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2)

  whw_lng <- cbind(whw_lng, col_split)
  names(whw_lng)[4] <- 'strike_mode'
  names(whw_lng)[5] <- 'city'
  whw_lng <- whw_lng %>% dplyr::select(-variable)
  whw_lng <- whw_lng %>% dplyr::filter(strike_mode != "rowSum")
  whw_lng <- whw_lng %>% rename(str_mode = mode, cas_mode = strike_mode)

  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode))

  qualified_scen_name <- 'Bus'
  scen <- 'Scenario'

  if (cs == 'bl'){
    qualified_scen_name <- 'Baseline'
    scen <- ''
  }else if(cs == "w_sc"){
    qualified_scen_name <- 'Walking'
  }else if(cs == "bi_sc"){
    qualified_scen_name <- 'Bicycle'
  }else if(cs == "car_sc"){
    qualified_scen_name <- 'Car'
  }else if(cs == "mc_sc"){
    qualified_scen_name <- 'Motorcyle'
  }

  readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))

  f <- ggplot(data = whw_lng) +
    aes(x = str_mode, fill = cas_mode, weight = value) +
    geom_bar(position = "stack") +
    scale_fill_brewer(palette = "Dark2") +
    labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
    coord_flip() +
    facet_wrap(vars(city), scales = "free_y")

  file_list[[cs]] <- f

  ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)

  d <- paste0("results/multi_city/whw_matrices/interactive_plots/whw_", cs,".html")

  htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
  #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")

  st[[cs]] <- format(td, scientific = F)

  print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', cs)))

}

```

### DT - Who hit whom matrix for all case cities {#whw_all}
Baseline

```{r dt}

#for (cs in names(all_whw)){
td <- NULL

cs <- 'Baseline'

var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()

fname <- paste("whw_", var, ".csv")

td <- st[[var]]

DT::datatable(td, 
              extensions = 'Buttons',
              escape = FALSE,
              rownames = FALSE,
              options = list(dom = 'Bfrtip',
                             buttons = 
                               list('colvis', list(
                                 extend = 'collection',
                                 buttons = list(list(extend='csv',
                                                     filename = fname),
                                                list(extend='excel',
                                                     filename = fname)),
                                 text = 'Download'
                               )),
                             scrollX = TRUE,
                             pageLength = nrow(td),
                             order=list(list(2,'desc')),
                             columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1))))))

# }

```

## TOTAL

Change in deaths total (for the city based on real population size) by age group by scenario

### Change in deaths {#change_death}

```{r scen_prop = "asis"}
death_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$deaths
    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat==z]))
  })
  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- scen_names
  temp
})

sum_and_round_and_print(death_totals,"Change in deaths total in ")
```

Change in deaths per 100,000 people by age group by scenario

### Change in deaths per 100,000 {#change_death_100k}

```{r scen_prop = "asis", echo = F, message = F}
death_rates <- lapply(cities,function(x) rbind(death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(death_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
names(death_rates) <- cities

round_and_print(death_rates,"Change in deaths per 100,000 in ")
```

YLLs total (for the city based on real population size) by age group  by scenario

<!-- ### Change in YLL {#change_YLL} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- yll_totals <- lapply(io,function(x){ -->
<!--   temp <- sapply(1:5,function(y){ -->
<!--     xx <- x$outcomes$hb$ylls -->
<!--     xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
<!--     sapply(sort(unique(xx$age_cat)),function(z) -->
<!--       sum(xxx[xx$age_cat==z])) -->
<!--   }) -->
<!--   colnames(temp) <- scen_names -->
<!--   temp -->
<!-- }) -->
<!-- sum_and_round_and_print(yll_totals,"Change in YLL total in ") -->
<!-- ``` -->

<!-- YLLs per 100,000 people by age group by scenario -->

<!-- ### Change in YLLs per 100,000 {#change_YLL_age_100k} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- yll_rates <- lapply(cities,function(x) rbind(yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(yll_rates) <- cities -->
<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
<!-- ``` -->

<!-- YLLs per 100,000 people by gender by age group scenario -->

<!-- ### Change in YLLs per 100,000 {#change_YLL_gender_age_100k} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- yll_totals <- lapply(io,function(x){ -->
<!--   temp <- sapply(1:5,function(y){ -->
<!--     xx <- x$outcomes$hb$ylls -->
<!--     rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
<!--   }) -->
<!--   rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2])) -->
<!--   colnames(temp) <- scen_names -->
<!--   temp -->
<!-- }) -->

<!-- yll_rates <- lapply(cities,function(x)  -->
<!--   rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/ -->
<!--           t(repmat(io[[x]]$demographic$population,5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(yll_rates) <- cities -->
<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
<!-- ```  -->

## BY PATHWAY

By pathway (with non- injury separate & summed)

Change in deaths total (for the city based on real population size) by age group by scenario

### Change in deaths due to injury {#change_death_injury}

```{r}
injury_totals <- lapply(io,function(x){
  xx <- x$outcomes$hb$deaths
  xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
  colnames(xxx) <- scen_names
  xxx
})
sum_and_round_and_print(injury_totals,"Change in deaths due to injury in ")
```

### Change in deaths due to injury per 100,000 km

```{r inj_100k = "asis", echo = F, message = F}
require(tibble)

overall_el <- list()

for (city in cities){

  #city <- 'delhi'

  print(city)

  el <- list()

  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){

  for (cs in names(io[[city]]$outcomes$whw)){
    # cs <- 'Baseline'
    # cs <- "Scenario 4"
    # city <- 'buenos_aires'
    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){

      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

    }

    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)

    if (length(el) == 0){
      el <- td4 %>% dplyr::select(stage_mode)
    }

    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')

    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()

    names(td4)[2] <- var

    td5 <- td4

    td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 100000, 4)
    # / sum(io[[city]]$demographic$population, na.rm = T)

    names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")

    el <- inner_join(el, td4, by = 'stage_mode')

    # for (i in 1:nrow(td4)){
    #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
    #   
    #   
    # }

    # print(td4)
  }
  print(kable(el, caption = city))

  td <- el %>% dplyr::select(-contains('count'))
  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")

  overall_el[[city]] <- td

  cat("\n")
  #kable(print(el), caption = city)
  #}

}
```

### Change in deaths due to injury per 100,000 km across cities

```{r inj_100k_all_cities = "asis"}
require(tibble)

td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()

td[is.na(td)] <- 0

td <- td %>% dplyr::select(stage_mode, sort(names(.)))

readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')

colnames(td) = gsub("bi_sc", "bisc", colnames(td))

colnames(td) = gsub("bus_sc", "bssc", colnames(td))

colnames(td) = gsub("car_sc", "csc", colnames(td))

colnames(td) = gsub("mc_sc", "mcsc", colnames(td))

colnames(td) = gsub("w_sc", "wsc", colnames(td))

injury_risks_lng <- reshape2::melt(td)

col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)

injury_risks_lng <- cbind(injury_risks_lng, col_split)
names(injury_risks_lng)[4] <- 'scenario'
names(injury_risks_lng)[5] <- 'city'
injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")

injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)

injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'

cols_city <- c("#e41a1c",
"#377eb8",
"#4daf4a",
"#984ea3",
"#ff7f00",
"#ffff33",
"#a65628",
"#f781bf",
"#999999")

cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')

d <- ggplot(injury_risks_lng) +
  aes(x = stage_mode, fill = scenario, weight = value) +
  geom_bar(position = "dodge", colour = 'black') +
  scale_color_manual(values = cols) +
  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
  coord_flip() +
  theme_minimal() +
  facet_wrap(vars(city)) +
  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per 100k km"))

ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_100k_km.png"), limitsize = F, dpi = 300, scale = 1.5)

fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_100k_km.html")

htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))

print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))

```

### Change in deaths due to injury per Billion km across cities

```{r inj_100k_all_cities = "asis"}
require(tibble)

injury_risks_b <- td

# Convert it into Billion by multiping the values by 10000
injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000

readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')

colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))

colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))

colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))

colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))

colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))

injury_risks_lng <- reshape2::melt(injury_risks_b)

col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)

injury_risks_lng <- cbind(injury_risks_lng, col_split)
names(injury_risks_lng)[4] <- 'scenario'
names(injury_risks_lng)[5] <- 'city'
injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")

injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)

injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'

temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))

d <- ggplot(injury_risks_lng) +
  aes(x = stage_mode, fill = scenario, weight = value) +
  geom_bar(position = "dodge", colour = 'black') +
  scale_color_manual(values = cols) +
  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
  coord_flip() +
  theme_minimal() +
  facet_wrap(vars(city)) +
  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km"))

df <- injury_risks_lng %>% filter(city != 'cape_town') %>% spread(value = value, key = city) 
df[df == 0] <- NA
df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
df1 <- df %>% dplyr::filter(variable == 'mv')
df <- df %>% dplyr::filter(variable != 'mv')

d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
   geom_bar(position = "dodge", colour = 'black') +
   scale_fill_manual(values = cols_cbrewer) +
   theme_minimal() + 
   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
   facet_wrap(vars(scenario)) + 
   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))

d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
    geom_bar(position = "dodge", colour = 'black') +
    scale_fill_manual(values = cols_cbrewer) +
    theme_minimal() + 
    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
    facet_grid(vars(scenario), scales='free_x') + 
    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))

ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)

fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")

htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))

ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')

fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")

htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))

print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))

fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")

htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))

```

### # of deaths due to injury per 100,000 people

```{r inj_100k = "asis"}
require(tibble)

overall_el_normalized <- list()

for (city in cities){

  print(city)

  el <- list()

  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){

  for (cs in names(io[[city]]$outcomes$whw)){
    # cs <- 'Baseline'
    # # cs <- "Baseline"
    # city <- 'accra'
    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){

      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

    }

    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)

    if (length(el) == 0){
      el <- td4 %>% dplyr::select(stage_mode)
    }

    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')

    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()

    names(td4)[2] <- var

    td5 <- td4

    td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)

    el <- inner_join(el, td4, by = 'stage_mode')
  }

  el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

  print(kable(el, caption = city))
  td <- el
  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")

  overall_el_normalized[[city]] <- td

  cat("\n")
  #kable(print(el), caption = city)
  #}

}
```

### Change in deaths due to PA {#change_death_PA}

```{r}

pa_totals <- lapply(io,function(x){
  xx <- x$outcomes$pathway_hb$deaths
  xxx <- sapply(1:5,function(y){
    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
  })
  colnames(xxx) <- scen_names
  xxx
})
sum_and_round_and_print(pa_totals,"Change in deaths due to PA in ")
```

### Change in deaths due to AP {#change_death_AP}

```{r}
ap_totals <- lapply(io,function(x){
  xx <- x$outcomes$pathway_hb$deaths
  xxx <- sapply(1:5,function(y){
    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
  })
  colnames(xxx) <- scen_names
  xxx
})
sum_and_round_and_print(ap_totals,"Change in deaths due to AP in ")
```

## Change in deaths per 100,000 people by age group by scenario

### Change in deaths due to injury per 100,000 {#change_death_injury_100k}

```{r scen_prop = "asis"}
injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
names(injury_rates) <- cities
round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
```

### Change in deaths due to PA per 100,000 {#change_death_PA_100k}
```{r}
pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
names(pa_rates) <- cities
round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")

```

### Change in deaths due to AP per 100,000 {#change_death_AP_100k}
```{r}

ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
names(ap_rates) <- cities
round_and_print(ap_rates,"Change in deaths due to AP per 100,000 in ")
```

YLLs total (for the city based on real population size) by age group by scenario

<!-- ### Change in YLLs due to injury {#change_yll_injury} -->

<!-- ```{r} -->
<!-- injury_totals <- lapply(io,function(x){ -->
<!--   xx <- x$outcomes$hb$ylls -->
<!--   xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y]))) -->
<!--   colnames(xxx) <- scen_names -->
<!--   xxx -->
<!-- }) -->
<!-- sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to PA {#change_yll_PA} -->
<!-- ```{r} -->
<!-- pa_totals <- lapply(io,function(x){ -->
<!--   xx <- x$outcomes$pathway_hb$ylls -->
<!--   xxx <- sapply(1:5,function(y){ -->
<!--     xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]]) -->
<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
<!--   }) -->
<!--   colnames(xxx) <- scen_names -->
<!--   xxx -->
<!-- }) -->
<!-- sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to AP {#change_yll_AP} -->
<!-- ```{r} -->
<!-- ap_totals <- lapply(io,function(x){ -->
<!--   xx <- x$outcomes$pathway_hb$ylls -->
<!--   xxx <- sapply(1:5,function(y){ -->
<!--     xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]]) -->
<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
<!--   }) -->
<!--   colnames(xxx) <- scen_names -->
<!--   xxx -->
<!-- }) -->
<!-- sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ") -->
<!-- ``` -->

<!-- YLLs per 100,000 people by age group by scenario -->

<!-- ### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(injury_rates) <- cities -->
<!-- round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k} -->
<!-- ```{r} -->
<!-- pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(pa_rates) <- cities -->
<!-- round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k} -->
<!-- ```{r} -->
<!-- ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(ap_rates) <- cities -->
<!-- round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ") -->
<!-- ``` -->

<!-- ## BY DISEASE -->

<!-- Change in deaths total (for the city based on real population size) by scenario -->

<!-- ### Change in deaths due to disease {#change_death_disease} -->

<!-- ```{r} -->

<!-- disease_totals <- lapply(io,function(x){ -->
<!--   xx <- x$outcomes$hb$deaths -->
<!--   xxx <- sapply(1:5,function(y){ -->
<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
<!--   }) -->
<!--   colnames(xxx) <- scen_names -->
<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
<!--   xxx -->
<!-- }) -->
<!-- sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ") -->

<!-- ``` -->

<!-- Change in deaths per 100,000 people by scenario -->

<!-- ### Change in deaths due to disease per 100,000 {#change_death_disease_100k} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(disease_rates) <- cities -->
<!-- round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ") -->
<!-- ``` -->

<!-- Change in YLL total (for the city based on real population size) by scenario -->

<!-- ### Change in YLL due to disease {#change_yll_disease} -->

<!-- ```{r} -->

<!-- disease_totals <- lapply(io,function(x){ -->
<!--   xx <- x$outcomes$hb$ylls -->
<!--   xxx <- sapply(1:5,function(y){ -->
<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
<!--   }) -->
<!--   colnames(xxx) <- scen_names -->
<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
<!--   xxx -->
<!-- }) -->
<!-- sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ") -->

<!-- ``` -->

<!-- ### Change in YLL due to disease per 100,000 {#change_yll_disease_100k} -->

<!-- ```{r scen_prop = "asis"} -->
<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
<!-- names(disease_rates) <- cities -->
<!-- round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ") -->
<!-- ``` -->