### 2020-11-02 code/trips_analysis/mode_speed_summary.Rmd
adding summary tables for distance

---
title: "mode_speed_summary"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

Summary table for mode speeds
```{r}
summary <- read.csv("../../data/analysis/mode_speed/speed_df.csv")
print(summary)
```

Boxplots for mode_speeds
```{r}
bangalore <- read.csv("../../data/local/bangalore/bangalore_trip.csv")
delhi <- read.csv("../../data/local/delhi/delhi_trip.csv")
vizag <- read.csv("../../data/local/vizag/vizag_trip.csv")
sao_paulo <- read.csv("../../data/local/sao_paulo/sao_paulo_trip.csv")

bangalore <- bangalore %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Bangalore",
         id = paste0("Bangalore",stage_id)) %>% 
  select(id, city, speed, stage_mode) %>% 
  rename(mode = stage_mode)

delhi <- delhi %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Delhi",
         id = paste0("Delhi", stage_id)) %>% 
  select(id, city, speed, stage_mode) %>% 
  rename(mode = stage_mode)

vizag <- vizag%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Vizag",
         id = paste0("Vizag", trip_id)) %>% 
  select(id, city, speed, trip_mode) %>% 
  rename(mode = trip_mode)

sao_paulo <- sao_paulo%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Sao_Paulo",
         id = paste0("Sao_Paulo", trip_id)) %>% 
  select(id, city, speed, trip_mode) %>% 
  rename(mode = trip_mode)

trip <- rbind(bangalore, delhi, vizag, sao_paulo)
trip_walk <- filter(trip, mode=="walk")
trip_bicycle <- trip[trip$mode=="bicycle",]
trip_bus <- trip[trip$mode == "bus",]
trip_car <- trip[trip$mode == "car",]

```

```{r}
boxplot(speed~mode, data = droplevels(trip), ylim=c(0,40))
```

## Walking
```{r}
boxplot(speed~mode+city, data = droplevels(trip_walk), ylim=c(0, 10))
```

## Cycling
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
```

## Bus
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bus), ylim=c(0, 40))
```

## Car
```{r}
boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
```