### 2020-11-09 code/trips_analysis/mode_speed_summary.Rmd

---
title: "mode_speed_summary"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(wesanderson)
```

## Notes

Bangalore had a beautiful variable called transfer time = walk time plus wait time, but when closely examined, what was recoreded there seemd to just be the travel time. So in our data processing, we had used both the variables on transfer time and start and stop time to come up with trip duration. Are there any concerns here?

## Summary table for mode speeds
```{r}
summary <- read.csv("../../data/analysis/mode_speed/speed_df.csv")
print(summary)
```

Boxplots for mode_speeds
```{r}
bangalore <- read.csv("../../data/local/bangalore/bangalore_trip.csv")
delhi <- read.csv("../../data/local/delhi/delhi_trip.csv")
vizag <- read.csv("../../data/local/vizag/vizag_trip.csv")
sao_paulo <- read.csv("../../data/local/sao_paulo/sao_paulo_trip.csv")

bangalore <- bangalore %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Bangalore",
         id = paste0("Bangalore",stage_id),
         age_group = ifelse(age<=50, "<50", "50+")) %>% 
  select(id, city, speed, sex, age_group ,stage_mode) %>% 
  rename(mode = stage_mode)

delhi <- delhi %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Delhi",
         id = paste0("Delhi",stage_id),
         age_group = ifelse(age<=50, "<50", "50+")) %>% 
  select(id, city, speed, sex, age_group ,stage_mode) %>% 
  rename(mode = stage_mode)

vizag <- vizag%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Vizag",
         id = paste0("Vizag", trip_id),
          age_group = ifelse(age<=50, "<50", "50+")) %>% 
  select(id, city, sex, age_group, speed, trip_mode) %>% 
  rename(mode = trip_mode)

sao_paulo_1 <- sao_paulo%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Sao_Paulo_1",
         id = paste0("Sao_Paulo_1", trip_id),
          age_group = ifelse(age<=50, "<50", "50+")) %>% 
  select(id, city,sex, age_group, speed, trip_mode) %>% 
  rename(mode = trip_mode)

sao_paulo_2 <- sao_paulo%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(trip_duration = (trip_duration - (walk_to_pt + walk_from_pt)),
         speed = trip_distance / ( trip_duration / 60),
         city = "Sao_Paulo_2",
         id = paste0("Sao_Paulo_2", trip_id),
         age_group = ifelse(age<=50, "<50", "50+")) %>% 
  select(id, city,sex, age_group,  speed, trip_mode) %>% 
  rename(mode = trip_mode)

trip <- rbind(bangalore, delhi, vizag, sao_paulo_1, sao_paulo_2)
trip_walk <- filter(trip, mode=="walk")
trip_bicycle <- trip[trip$mode=="bicycle",]
trip_motorcylce <- trip[trip$mode=="motcycle",]
trip_bus <- trip[trip$mode == "bus",]
trip_car <- trip[trip$mode == "car",]

```

```{r}
boxplot(speed~mode, data = droplevels(trip), ylim=c(0,40))
```

## Walking
```{r}
boxplot(speed~mode+city, data = droplevels(trip_walk), ylim=c(0, 10))
```
## Walking by sex and age
```{r}
trip_walk %>% 
  filter(!is.na(sex) & !is.na(age_group)) %>% 
  ggplot(aes(x = city, y = speed,  fill = sex))+
  geom_boxplot(aes(col = age_group)) + 
  ylim(0, 8)+
  scale_fill_manual(values= c("grey80", "bisque2"))

```

## Cycling
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
```

## Cycling by sex and age
```{r}
trip_bicycle %>% 
  filter(!is.na(sex) & !is.na(age_group)) %>% 
  ggplot(aes(x = city, y = speed,  fill = sex))+
  geom_boxplot(aes(col = age_group)) + 
  ylim(0, 15)+
  scale_fill_manual(values= c("grey80", "bisque2"))

#trip_bicycle %>% filter(city=="Delhi") %>% View
```

## Motorcycling
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
```

## Car
```{r}
boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
```

## Bus
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bus), ylim=c(0, 40))
```

## Walk to/from PT
```{r}
walk_pt <- sao_paulo %>% 
  select(walk_to_pt, walk_from_pt) %>% 
  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
  filter(walk_to_from_pt > 0)

walk_pt <- melt(walk_pt)

boxplot(value~variable, data = droplevels(walk_pt), ylim=c(0, 20))
```