### 2021-05-04 code/producing_trips_rd/trip-summary.Rmd
Regen travel survey summary

---
title: "Baseline Travel Survey Summary ITHIM-R"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: true
---

```{r, include = FALSE}
source("used_functions.R")
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
```

```{r loadLibraries, echo = F, message = F}

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
                                            # much better results. Your mileage may vary

```

```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
# (readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
#   ) %>% kbl() %>% kable_classic()

# (If short walk is not explicitly recorded, we have to find a way to remove; if they are missing, what should we do?)

# Sao Paulo:
# 1) Walking trips: Page 60, Pesquisa de Mobilidade da RMSP –2012 –Manual da Pesquisa Domiciliar
# Belo horizonte:
# 1) Study area: page 35, codebook and protocol.pdf
# Cape town:
# 1) Main mode Hierarchy: page 45, Final Draft HH Survey Report.pdf
```

```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Buenos_Aires <- value

trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Sao_Paulo <- value

trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Belo_Horizonte <- value

trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Mexico <- value

trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Bogota <- value

trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Santiago <- value

trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Visakhapatnam <- value

trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Delhi <- value

trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Bangalore <- value

trip <- read_csv("../../data/local/accra/accra_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Accra <- value

trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
                 guess_max = 100000)
quality_check(trip)
trip_summary$Cape_Town <- value
```

```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Buenos_Aires <- value

trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Sao_Paulo <- value

trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Belo_Horizonte <- value

trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Mexico <- value

trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Bogota <- value

trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Santiago <- value

trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Visakhapatnam <- value

trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Delhi <- value

trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Bangalore <- value

trip <- read_csv("../../data/local/accra/accra_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Accra <- value

trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
                 guess_max = 100000)
quality_check(trip, 15, 69)
trip_summary2$Cape_Town <- value
```

## summary table (all rows)
```{r, include = TRUE,  echo= FALSE, results='asis'}
kable(trip_summary) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14, fixed_thead = T)

```

## summary table (only people between 15 and 69)
```{r, include = TRUE,  echo= FALSE, results='asis'}
kable(trip_summary2) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14, fixed_thead = T)

```