### 2021-08-25 latex/output-md.Rmd

# Introduction

## Summary of literature

## Summary of aims

## Overview

The remainder of the document presents a current working version of the
ITHIM, and is a comprehensive record of all the calculations of which it
is composed. The equations for the ITHIM are described in Section
[2](#equations){reference-type="ref" reference="equations"} (tabulated
in Section [7](#tableequations){reference-type="ref"
reference="tableequations"}). The method is concretised through the
example application to Accra in Section
[3](#example){reference-type="ref" reference="example"} and particular
columns in Tables [1](#indices){reference-type="ref"
reference="indices"} and [3](#inputs){reference-type="ref"
reference="inputs"}, with results presented in Section
[19](#results){reference-type="ref" reference="results"}.

The ITHIM as described here is implemented in the ITHIM-R R package
(<https://github.com/ITHIM/ITHIM-R>). Other cities are also included in
that package (São Paulo, Delhi, Bangalore), and reference will be made
to those where relevant.

# Method: ITHIM {#equations}

In the ITHIM, we combine many information inputs and compute the health
burden through three pathways (depicted in Figure
[1](#workflow){reference-type="ref" reference="workflow"}). Each pathway
has its own set of definitions and outputs, and we refer to them as
"modules".

![ITHIM workflow. Inputs to the model have a bold, black outline. The
graph depicts parent--child relationships, where a child (at the head of
an arrow) depends upon all its parents (at the source(s) of the
arrow(s)). To aid visualisation, an arrow connected to the outside of a
grey box indicates that the arrow connects to each item within the box.
*AT: active travel. MET: metabolic equivalent task. pp: per person. AP:
air pollution. PA: physical activity. DR: dose--response relationship.
GBD: global burden of disease. BOD: (local) burden of
disease.*](flowchart.pdf){#workflow width="90%"}

To describe the model, we use lower-case letters to denote indices, or
dimensions, of objects. They take one of a set of possible values,
detailed in Table [1](#indices){reference-type="ref"
reference="indices"}. The set of input items are denoted by capital
letters, and variable parameters are denoted by Greek letters. These are
presented in Table [3](#inputs){reference-type="ref"
reference="inputs"}. The equations of the ITHIM are presented below, and
summarised in Table [4](#calculationstab){reference-type="ref"
reference="calculationstab"}. These use letter modifiers to show how the
inputs are modified in the course of output generation.

We calculate the ITHIM by combining the modules' outputs in terms of the
health burden ($\hat{U}_{a,g,h,o,s}$). Other outputs of the model
include: MMETs per person ($M_{i,s}$), PM2.5 in each scenario
($\bar{P}_s$), PM2.5 per person ($\check{W}_{i,s}$), and injuries burden
($\check{I}_{a,g,m,o,s}$).

To assess uncertainty, we sample from the ITHIM output by sampling
uncertain parameters multiple times and calculating the ITHIM. We
specify parametric distributions or sampling strategies for all the
uncertain parameters. Evaluating the ITHIM with uncertain parameters
allows assessment of their impact on the outcome (AKA sensitivity
analysis). We use EVPPI to calculate the expected reduction in
uncertainty in the outcome were we to learn a parameter perfectly. This
means we can implement models that are basic in their parametrisation,
and learn at the end for which parameters it would be worthwhile
spending dedicated time learning better.

::: {.center}
::: {#indices}
  Label   Name                   Levels
  ------- ---------------------- ---------------------
  $a$     Age group              15--49
                                 50--69
  $g$     Gender                 Male
                                 Female
  $h$     Disease                
  $i$     Individual index       
  $j$     Trip index             
  $m$     Transport mode         walk
                                 cycling (cyc.)
                                 bus
                                 car
                                 motorbike (mot.)
                                 goods vehicles (GV)
                                 subway (sub.)
  $o$     Outcome                Death
                                 YLL
  $s$     Scenario               
  $w$     AP RR hyperparameter   $1,2,3,4$
  $x$     PA RR hyperparameter   $1,2,3$
  $z$     PA dose                $0,...,300$

  : Indices used in the ITHIM variables. The "Label" is the subscript
  used to refer to the index in this document. The "Name" is
  descriptive. The "Levels" are the levels used in the Accra model.
:::
:::

::: {.center}
::: {#abbrev}
  Abbreviation   Meaning
  -------------- -----------------------------------------------
  AP             Air pollution
  CDF            Cumulative distribution function
  COPD           Chronic obstructive pulmonary disease
  DR             Dose response
  EVPPI          Expected value of perfect partial information
  GAM            Generalised additive model
  GBD            Global burden of disease
  IHD            Ischemic heart disease
  ITHIM          Integrated transport and health impact model
  LRI            Lower respiratory infection
  MMET           Marginal metabolic equivalent task
  PA             Physical activity
  PAF            Population attributable fraction
  PIF            Potential impact fraction
  PM2.5          Particulate matter (diameter $< 2.5 \mu$m)
  PT             Public transport
  RR             Relative risk
  VOI            Value of information
  YLL            Years of life lost

  : Common abbreviations.
:::
:::

::: {.landscape}
::: {.center}
::: {#inputs}
  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------
  [\[inputs\]]{#inputs label="inputs"}                                                                          *ITHIM-R reference*                                                
  Name                                        Description                                                       Value in Accra model                                         Label   Model/ Setting
  *Global values*                                                                                                                                                                  
  $U_{a,g,h,o}$                               Background burden of disease                                      Constant                                                  GBD_DATA          Setting
  $N_{a,g}$                                   Population by age and gender in GBD_DATA                          Constant                                                  GBD_DATA          Setting
  $\bar{N}_{a,g}$                             Population by age and gender                                      Constant                                               DEMOGRAPHIC          Setting
  $\rho_h$                                    Chronic disease scalar                                            1 or Lnorm(0,0.18)                          CHRONIC_DISEASE_SCALAR          Setting
  *Travel values*                                                                                                                                                                  
  $Z_{m=walk}$                                Walk speed                                                        4.8                                                    MODE_SPEEDS          Setting
  $Z_{m=cyc.}$                                Cycle speed                                                       14.5                                                   MODE_SPEEDS          Setting
  $Z_{m=bus}$                                 Bus speed                                                         15                                                     MODE_SPEEDS          Setting
  $Z_{m=car}$                                 Car speed                                                         21                                                     MODE_SPEEDS          Setting
  $Z_{m=mot.}$                                Motorbike speed                                                   25                                                     MODE_SPEEDS          Setting
  $\epsilon$                                  Walk time to bus                                                  5 or Lnorm(log(5),0.18)                              BUS_WALK_TIME          Setting
  $\mu_{m=HGV}$                               Truck distance relative to car                                    0.21 or Beta(3,10)                              TRUCK_TO_CAR_RATIO          Setting
  $\mu_{m=bus}$                               Bus driver distance relative to bus passenger distance            1/45 or Beta(20,600)                        BUS_TO_PASSENGER_RATIO          Setting
  $\mu_{m=mot.}$                              Motorcycle fleet distance relative to total motorcycle distance   37/68 or Beta(37/3+1,31/3+1)             FLEET_TO_MOTORCYCLE_RATIO          Setting
  $T_j$                                       Trip-level data                                                   Constant                                           SYNTHETIC_TRIPS          Setting
  *Injury model values*                                                                                                                                                            
  $I_{a,g,m_{\text{vic}},m_{\text{str}}}$     Fatalities table                                                  Constant                                              INJURY_TABLE          Setting
  $\sigma$                                    Injury reporting rate                                             1 or Beta(8,3)                               INJURY_REPORTING_RATE          Setting
  $\omega$                                    Sum of distance exponents                                         1.9 or Lnorm(log(1.9),log(1.03))                  SIN_EXPONENT_SUM            Model
  $\psi$                                      Casualty fraction of $\omega$                                     0.5 or Beta(20,20)                    CASUALTY_EXPONENT\_ FRACTION            Model
  *Pollution model values*                                                                                                                                                         
  $\eta$                                      Background PM2.5 concentration                                    50 or Lnorm(log(50),0.18)                             PM_CONC_BASE          Setting
  $\zeta$                                     Fraction of measured PM2.5 concentration due to road transport    0.225 or Beta(5,20)                                 PM_TRANS_SHARE          Setting
  $\gamma_{m}$                                Vehicle emission inventory                                        Constant or Dirichlet                           EMISSION_INVENTORY          Setting
  $P$                                         Vehicle emission confidence                                       1                                    EMISSION_INVENTORY_CONFIDENCE          Setting
  *Pollution & health model values*                                                                                                                                                
  $C_1$                                       Base-level inhalation rate                                        1                                     BASE_LEVEL_INHALATION \_RATE            Model
  $C_2$                                       Exposure ratio window closed                                      0.5                                       CLOSED_WINDOW_PM\_ RATIO            Model
  $C_3$                                       Proportion of travel with closed windows                          0.5                                            CLOSED_WINDOW_RATIO          Setting
  $C_4$                                       Parameter for on-road PM2.5 level                                 3.216                                               ROAD_RATIO_MAX            Model
  $C_5$                                       Parameter for on-road PM2.5 level                                 0.379                                             ROAD_RATIO_SLOPE            Model
  $C_6$                                       Exposure ratio for underground                                    0.8                                                SUBWAY_PM_RATIO          Setting
  $H_{a,h,w}$                                 AP dose--response--curve parameters                               Constant                                                     DR_AP            Model
  $\xi_{h,w}$                                 Quantiles for AP RR functions                                     0.5 or Unif(0,1)                       AP_DOSE_RESPONSE\_ QUANTILE            Model
  *Physical activity model values*                                                                                                                                                 
  $\chi$                                      Day-to-week travel scalar                                         7 or 7$\times$Beta(20,3)               DAY_TO_WEEK_TRAVEL\_ SCALAR            Model
  $Y_i$                                       Individual non-transport MMETs                                    Constant                                PA_SET\$work_ltpa_marg_met          Setting
  $\theta$                                    Background PA scalar                                              1 or Lnorm(0,0.18)                            BACKGROUND_PA_SCALAR          Setting
  $\tilde{Y}$                                 Confidence in PA survey                                           1                                         BACKGROUND_PA_CONFIDENCE          Setting
  $\tilde{\theta}$                            Background PA zeros quantile                                      0.5 or Unif(0,1)                               BACKGROUND_PA_ZEROS          Setting
  $\lambda_{m=\text{cyc.}}$                   Cycling MET surplus                                               4.63 or Lnorm(log(4.63),1)                            MMET_CYCLING            Model
  $\lambda_{m=\text{walk}}$                   Walking MET surplus                                               2.53 or Lnorm(log(2.53),1)                            MMET_WALKING            Model
  $\lambda_{m\not\in\text{walk,cyc.}}$        In-vehicle MET surplus                                            0                                                                             Model
  *Physical activity & health model values*                                                                                                                                        
  $G_{h,x,z}$                                 Look-up table for relative risk truncated normal distribution     Constant                                                                      Model
  $\phi_{h}$                                  Quantiles for PA RR functions                                     0.5 or Unif(0,1)                       PA_DOSE_RESPONSE\_ QUANTILE            Model
  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------

  : All values input to the ITHIM model.
:::
:::
:::

## Data module {#data}

Input data for the ITHIM are listed as Roman capital letters in Table
[3](#inputs){reference-type="ref" reference="inputs"}. They include
(vehicle) mode speeds, injury/fatality records, disease dose--response
relationships, measures of air pollution, surveys of travel and physical
activity, population numbers by demographic group, MMETs associated with
active travel, and current burden of disease (see Figure
[1](#workflow){reference-type="ref" reference="workflow"}).

### Synthetic {#synth}

We create a "synthetic population" by matching individuals surveyed for
travel to the individuals surveyed for physical activity, based on age
and gender. Then each person in the synthetic population has a set of
trips taken and an amount of non-travel physical activity. This
population is assumed to be representative of the population of the city
or setting under study.

#### Alternatively/additionally

we could create a synthetic population as described in Appendix
[8.3](#synthpop){reference-type="ref" reference="synthpop"}, which
allows uncertainty in the propensity to travel. This method operates
with person-level, rather than trip-level, information.

### Scenarios

From the synthetic population we create "scenarios" of different
pictures of travel by changing certain trips (e.g. swapping the mode of
travel for particular trips and/or particular people). For the purposes
of illustration, we choose a simple scenario for Accra in which every
person gains one one-kilometre walk.

### Distances

For the ITHIM, we process the travel data into distance data of various
formats. First, we augment PT trips with a walk component: we take the
set $\{T_j:m(j)=\text{bus}\}$, which has $J$ entries. We add $J$
journeys with mode "walk" and duration $\epsilon$ to the set $T$. (This
step would not be necessary for a travel survey which already includes
all stages for each trip.)

Then we summarise the total distances and durations. The distance set is
labelled $\hat{T}_{m,s}$, representing the total distance travelled per
mode per scenario, and the duration set is constructed analogously. We
make an assumption that the distance for bus drivers scales linearly
with bus passengers, where the distance in the baseline is defined based
on the bus driver distance relative to car.

Distance and duration summaries include all person travel in the
synthetic population and, in addition, some trips that are attributed to
no-one in the synthetic population. These trips are used to represent
"fleet" travel, which is uncaptured in travel surveys, and unchanged in
scenarios. It's important to include these trips because for modes such
as motorcycle, where a considerable component of travel comes from
fleet, attributing all the harms associated with the mode to only those
surveyed dramatically overestimates the contribution of the individuals.
In scenarios, then, we would be predicting emissions and injury rates
that are far too high per person.

## AP module

### Background PM2.5 {#dir}

We use the total distance by mode ($\hat{T}_{m,s}$) to calculate the
total PM2.5 in the scenario ($\bar{P}_s$). First we calculate the
emission inventory in terms of fractions. If the confidence $P$ is 1
then we use $\dot{P}_{m}=\gamma_m$ directly. If $P<1$, we sample the
emission inventory fractions as follows:
$$\tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)),$$
$$\dot{P}_m=\tilde{\gamma}_m$$ where we choose some mapping function
such as $f(P,\gamma_m)=\frac{\gamma_m}{\sum_m\gamma_m}10^{5P}$ (Figure
[5](#emission_dist){reference-type="ref" reference="emission_dist"}).

Then we multiply vehicle distance by vehicle emission factor (where the
emission factor is the emission inventory (fraction) divided by distance
at baseline).
$$\tilde{P}_{m,s}=\dot{P}_m\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}.$$

For the baseline, $\tilde{P}_{m,s}$ is equal to the emission inventory
fraction. For each scenario, it is scaled. We take the sum over modes to
get a scalar for emissions in scenarios.
$$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}.$$ Then the background PM2.5 in
each scenario is the sum of the transport component and the
non-transport component: $$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta).$$

### PM2.5 per person

Dose--response relationships are based on background concentrations in
cities, and not on personal exposures. In our person-level model, we
expect that individuals will have different exposures based on their
different travel behaviours. While we can neither estimate individual
exposures nor map them via any dose--response relationship to a relative
risk, we aim to capture interpersonal differences via assumptions about
the mechanistic relationships between behaviours and exposures, and
anchor those to the background air pollution in their locality, for
which dose--response estimates of relative risk exist.

Therefore, when we write "individual exposure", it should not be read as
an aetiologically relevant exposure, but rather it should be read as a
unit- and scale-agnostic metric that maps an individual to a particular
dose with reference to city-level measurements of air pollution. New
research is required to elicit dose--response relationships between
individual health exposures and health outcomes \[citation from JPO?\].
Our approximations act as assumptions for now and placeholders until
more detailed data are available. \[Perhaps we should use words other
than "individual exposure". How about "individual PM2.5 reference
values"?\]

Individual exposures to PM2.5 are calculated using the background PM2.5
($\bar{P}_s$) and the trip sets. There are three major components to
daily exposure: one, a person's total inhalation off road; two, a
person's inhalation on road in a vehicle, and, three, a person's
inhalation on road while undertaking active transport. Each category has
an amount of background PM2.5 and a ventilation rate which together
inform overall exposure.

The ratio of exposure off road to that on road is a function of total
PM2.5, defined as $$K_s=C_4-C_5\log(\bar{P_s})$$ as in @Goel2015. This
defines the exposure of a person in an open vehicle (i.e. pedestrian,
cyclist or motorist):
$$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$$ and it is used to
calculate in-vehicle exposure, assuming an exposure of $C_2$ with the
window closed, and a proportion $C_3$ of vehicles having closed windows:
$$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3).$$
The exposure in a subway is constant and not dependent on the road ratio
$K_s$: $$\tilde{K}_{m=\text{sub.},s}=C_6.$$

Ventilation rates are calculated for each mode, assuming a base-level
inhalation rate of $C_1$: $$V_m=C_1+\frac{1}{2}\lambda_m,$$ where
$\lambda_m$ is the MMETs for mode $m$.[^1] Then the air inhaled during
travel per person is
$$\tilde{V}_{i,m,s} = \smashoperator{\sum_{\substack{j:i(j)=i,\\m(j)=m,\\s(j)=s}}}T_j\cdot V_{m(j)}/Z_{m(j)}$$
and the rate of inhalation during travel is
$$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}.$$ The
air inhaled when not travelling is
$$\hat{V}_{i,s} = C_1\left(24-\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}T_j/Z_{m(j)}\right).$$
Together, the PM2.5 exposure is calculated as the total PM2.5 inhaled
per hour as follows:
$$\check{W}_{i,s} = \frac{\bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})}{24}.$$

### AP--disease dose--response relative risk

We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate
their relative risk of five diseases (IHD, lung cancer, COPD, stroke,
LRI), using curves parametrised by four disease-specific variables
[@Burnett2014]. Of the five diseases, two (IHD and stroke) have
parameters specific to age groups starting at age 25.[^2] The other
three (lung cancer, LRI and COPD) have one set of parameters for all
ages.

The curves are in the form of samples of the set of four parameters. We
model the densities of these samples (using a quantile for parameter 3,
kernel density estimation for parameter 2, and GAMs for parameters 1 and
4)[^3] in order to draw either the median or random samples via their
quantiles ($\xi_{h,w}$) as follows:

$$\begin{aligned}
\tilde{H}_{a,h,w=3}=&\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})\\
\tilde{H}_{a,h,w=2}=&\text{CDF}_{H_{a,h,w=2}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=2})\\
\tilde{H}_{a,h,w=1}=&\text{CDF}_{H_{a,h,w=1}|\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=1})\\
\tilde{H}_{a,h,w=4}=&\text{CDF}_{H_{a,h,w=4}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=1}}^{-1}(\xi_{h,w=4}).\end{aligned}$$

From these parameters, the relative risk of mortality is defined
$$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=1}\times
\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=2}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=4}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}.$$
For $h\not\in\{\text{IHD, lung cancer, COPD, stroke, LRI}\}$,
$\check{H}_{h,i,s} \equiv 1$.

## PA module

### Individual-level MMETs {#beta}

Using trip sets and the synthetic population, we calculate total MMETs
per person ($M_{i,s}$) as the sum of walking MMETs per day and cycling
MMETs per day, which are scaled up to a week via the scalar $\chi$, and
work/leisure MMETs per week as follows:
$$M_{i,s}=\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}
\chi\lambda_{m(j)}\frac{T_j}{Z_{m(j)}}+\theta Y_i.$$

Here, $\lambda_m$ is the MMETs for mode $m$, $T_j$ the distance of trip
$j$, $Z_m$ the speed of mode $m$, $\theta$ the scalar for background PA,
and $\check{Y}_i$ the amount of work and leisure (background) PA of
person $i$.

We calculate $\check{Y}_i$ according to the confidence in the PA survey,
$\tilde{Y}$. First, we calculate the raw probability that a person in
demographic group completes no non-travel PA:
$$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}.$$ Then, we set the
probability to use, $\bar{Y}_{a,g}$, as $\dot{Y}_{a,g}$ if our
confidence $\tilde{Y}$ is 1. If $\tilde{Y}<1$, we map the confidence to
parametrise a Beta distribution somehow, e.g. (Figure
[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"})
$$\begin{aligned}
\hat{Y} &= 500^{\tilde{Y}+0.2},\\
\beta & = \hat{Y}\dot{Y}_{a,g}\left(\frac{1}{\dot{Y}_{a,g}} -1\right),\\
\alpha & = \hat{Y} - \beta,\\
\bar{Y}_{a,g} & = \text{CDF}_{\text{Beta}(\alpha,\beta)}^{-1}(\tilde{\theta}).\end{aligned}$$

Finally, for each person in the population, we sample non-travel MMETs
as zero with probability $\bar{Y}_{a,g}$ and from the raw non-zero
density of their demographic group with probability $1-\bar{Y}_{a,g}$.

### PA--disease dose--response relative risk

We use each person's MMETs per week (${M}_{i,s}$) to calculate their
relative risk of six diseases, using curves found from meta analysis.
Each disease (except type 2 diabetes) has a threshold beyond which there
is no further change in relative risk. (Total cancer, breast cancer,
colon & rectum cancer, endometrial cancer, & coronary heart disease: 35;
lung cancer: 10; stroke: 32; all cause: 16.08.) For all individual
cancers, we estimate the burden in terms of . For all other diseases, we
use mortality.

The interpolation of the dose--response curve can be described as
follows:
$$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$$
where we interpolate the mean, the lower bound, and the upper bound.
These define the normal from which we sample, truncated at 0:
$$F_S(v_x) = \mathcal{N}\left.\left(v_{x=2},\frac{v_{x=3}-v_{x=1}}{1.96}\right)\right|_{F_S(v_x)>0}.$$

Then the relative risk for each person for each disease is calculated
for the quantile $\phi_h$ as
$$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right).$$
See Figure [2](#cancer){reference-type="ref" reference="cancer"} for an
example. For $h\not\in$ {diabetes, total cancer, breast cancer, colon &
rectum cancer, endometrial cancer, IHD, lung cancer, stroke, all cause},
$W_{h,i,s} \equiv 1$.

![Example of the dose--response workflow for PA and total cancer. Top
left: results of the metaanalysis: mean, lower bound and upper bound of
the relationship between PA dose and relative risk of disease. Top
right: examples of 500 samples from the normal distribution defined by
the mean of the top-left panel and the standard deviation defined by
(upper bound minus lower bound) divided by 1.96 and truncated at 0.
Bottom left: relative risks of individuals with PA doses 5, 25 and 45
are found by mapping their doses onto the median curve. This is used in
the "constant" ITHIM use case. Bottom right: relative risks of
individuals with PA doses 5, 25 and 45 for three different samples from
the distribution shown in the top-right panel. For each sample, each
individual is mapped onto the response curve it defines.
](cancer_PA_dose_response.pdf){#cancer width="60%"}

## Injury module

### Processing

For the injury module, we require distance per travel mode per
demographic group per scenario. It requires matching of mode names in
the travel survey to mode names in the injury records, for example
"taxi", "shared auto", "shared taxi" and "car" combine to form "car".
Some work is required to separate drivers from passengers; we currently
assume all travellers are drivers, with the exception of "bus".[^4]

### Modelling

We model injuries via regression by predicting the number of fatalities
of each demographic group on each mode, which we calculate as a sum over
all the ways in which they might have been injured (i.e., all the modes
with which they might have collided).

The predictive covariates include the distances travelled by the parties
and their demographic details. These requirements lead to a natural
separation of the (training) dataset into two groups: a set for which we
have distance data for the other party, and a set for which we do not
have distance data for the other party. The former equates to a "who hit
whom" ("whw") matrix (albeit in a higher dimension), and will account
for changes in injuries resulting of a change in strike-mode travel. The
latter corresponds to causes of injury that will not change across
scenarios, including "no other vehicle" and modes of transport that we
do not consider to change, which might include trucks and buses if they
are not somehow explicitly included in the trip set. We label this group
"noov": no or other vehicle.

We model the number of injuries as a Poisson-distributed variable with
an offset depending on the distance(s) and the reporting rate. We write
the model as follows: $$\begin{aligned}
\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s,y} \sim &
\text{Poisson}(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}); \\
% 'count~
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{whw}}\}\right) = &
% cas_mode*strike_mode-log(injury_reporting_rate)+log(weight)
{\kappa_0} +\sum_{i=1}^n\kappa_iX_i-\log(\sigma)+\nonumber\\
% +log(cas_distance)
 & \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})
 %+log(strike_distance)+(STR_EXPONENT-1)*log(strike_distance_sum)
 +E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s});\\
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{noov}}\}\right) =&
{\kappa'_0}
+\sum_{i=1}^n\kappa'_iX'_i 
-\log(\sigma)
% +log(cas_distance)
+ \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s}).\end{aligned}$$
We choose this form of equation to enable (a) linearity in injuries with
respect to total travel combined across modes and (b) linearity in
injuries as subdivided by travellers of each mode.[^5] Note that this
methodology is under development. See
<https://github.com/robj411/safety_in_numbers_power_law>.

In general, one makes the best model that one can with the data
available. We use, for Accra, the covariates
`casualty mode*strike mode, casualty age,` and `casualty gender`, to
form the model matrix $X$. We also have another covariate, `year`, but
our travel data are for one year only, so we model instead many
observations of the same quantities.

We fit the coefficients $\kappa$ of the model using data for ten years,
assuming the same travel each year, which corresponds to the scenario
$s=\text{baseline}$. To incorporate the injury reporting rate
($\sigma$), we set it as an offset, whose value is 1 in fitting the
model, and as specified in making predictions.

### Prediction

We predict the number of injuries in each scenario based on the training
model built from the baseline scenario. For Accra, we predict for the
year 2016, using scenario-specific travel data. The number of fatalities
is taken as the sum of fatalities over the strike modes:
$$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}.$$
We extrapolate injury deaths to injury YLLs via the ratio in the GBD:
$$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}},$$
$$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}.$$

## Health module

### PA and AP relative risks combined

We combine the relative risks of disease as a function of AP and disease
as a function of PA through multiplication:
$$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}.$$ Not all diseases
have a dose--response relationship for both AP and PA. Just stroke, lung
cancer, and IHD have both. For the other diseases, only one of
$W_{h,i,s}$ and $\check{H}_{h,i,s}$ is different from one.

### Total disease burden

We calculate the total health burden relative to the reference scenario
(which, for our Accra example, is the baseline) using the injury and
health outputs combined with GBD data, via population attributable
fractions (PAF). The PAF is defined:
$$\hat{W}_{a,g,h,s}=\smashoperator{\sum_{\substack{i:a(i)=a,\\g(i)=g}}}\tilde{W}_{h,i,s}.$$

We then calculate the PAF relative to the reference scenario ("ref") as:
$$\bar{W}_{a,g,h,s}= \frac{\hat{W}_{a,g,h,s=\text{ref}}-\hat{W}_{a,g,h,s}}{\hat{W}_{a,g,h,s=\text{ref}}}.$$

We estimate the background burden of disease using GBD data and scaling
based on the ratio of populations between country and city. For country
populations $N_{a,g}$, city populations $\bar{N}_{a,g}$ and country
burden $U_{a,g,h,o}$, we estimate city burden as
$$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}.$$

If we are scaling the background burden of non-communicable diseases
("Neoplasms", "Ischemic heart disease", "Tracheal, bronchus, and lung
cancer", "Breast cancer", "Colon and rectum cancer", "Uterine cancer"),
we do so here. (Otherwise set $\rho\equiv 1$.)
$$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}.$$

We combine the burden with the PAF:
$$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}.$$

### Injury burden

And, finally, for the injuries, we sum over modes to compute the burden,
$$\bar{U}_{a,g,o,s}=\sum_m\check{I}_{a,g,m,o,s},$$ and subtract from the
values for the reference scenario:
$$\hat{U}_{a,g,h=\text{road injury},o,s}=\bar{U}_{a,g,o,s=\text{ref}}-\bar{U}_{a,g,o,s}.$$

# Example: Accra {#example}

Here we show the ITHIM as evaluated for the city of Accra, with a simple
scenario of every person in the synthetic population gaining a single, 1
km walk. We define XX uncertain univariate parameters (Section
[3.1](#uni){reference-type="ref" reference="uni"}), two confidences, and
two uncertain parameter sets (Section [3.3](#dr){reference-type="ref"
reference="dr"}).

## Parametric distributions for uncertain variables {#uni}

The XX uncertain univariate parameters are shown in Figure
[3](#distributions){reference-type="ref" reference="distributions"}.
Most of the parameters are self explanatory. Cycling and walking MMETs
are the number of MMETs per hour when undertaking cycling and walking,
and determine also the ventilation rates. Motorcycle distance is the
total distance travelled by motorcycles relative to the total distance
travelled by cars in the baseline scenario. Non-travel PA, injury
reporting rate and NCD burden all act as scalars for the corresponding
datasets. Note that the non-travel PA scalar does not affect the
$\approx$`<!-- -->`{=html}40% of the population whose non-travel PA is
0.

![Distributions of univariate uncertain parameters (1024
samples).](uncertain_parameters.pdf){#distributions width="95%"}

## Confidences

For physical activity propensity and emission inventories, we use
"confidences", values between 0 and 1 that represent how confident we
are about the data source. We map these values to Beta (Figure
[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"}) and
Dirichlet (Figure [5](#emission_dist){reference-type="ref"
reference="emission_dist"}) distributions, respectively. The
Beta-distributed propensity towards non-travel physical activity is
described in Section [2.3.1](#beta){reference-type="ref"
reference="beta"}. The Dirichlet-distributed emissions inventory is
described in Section [2.2.1](#dir){reference-type="ref"
reference="dir"}.

![Distributions of a Beta-distributed propensity to non-travel PA with
varying confidence values in the raw input data. The sampled value gives
the fraction of the population that engages in non-travel PA.
](prob_zero_PA.pdf){#prob_zero_PA width="50%"}

![Distributions of Dirichlet-distributed emission inventories for four
modes and varying confidence values. Raw input values were 4, 4, 20, and
60. ](emission_dist.pdf){#emission_dist width="65%"}

## Dose--response relationships {#dr}

For the dose--response relationships between physical activity (PA) and
disease and air pollution (AP) and disease, we assume that there is
uncertainty, but no variability, in the relationship. This means that we
sample a relationship from the distribution of relationships, and apply
that relationship to all individuals precisely. This means that, given
fixed doses, responses between individuals will be perfectly correlated.

We achieve this by use of the probability integral transform: we sample
a random variable uniformly distributed on the space (0,1) and map it,
via a cumulative distribution function, to the distribution describing
the dose--response relationship. See Figure
[6](#cdf){reference-type="ref" reference="cdf"} for an illustration.

![We use the same uniform samples to generate random variables from
different distributions. It means that the samples are perfectly
correlated.](DR_transform.pdf){#cdf width="80%"}

### Physical activity

Each disease's PA dose--response relationship is defined by a truncated
normal. For each dose, there is a mean value, an upper bound, and a
lower bound. For each person's dose, we get the response by mapping the
uniform random variable onto the truncated normal defined by the mean
and bounds for that dosage.

### Air pollution

For the AP relationship, there are four parameters per disease. We
sample the first from an empirical distribution using the probability
integral transform. We sample the second via the same method,
conditioned on the value of the first, constructing their joint density
with e.g. kde2d. The third parameter is sampled conditioned on the first
and second, constructing their joint density using a GAM. The final
parameter is sampled conditioned on the first, second, and third,
constructing their joint density using a GAM.

As before, there is perfect correlation between individuals, i.e. if
person A's dose is greater than person B's, then person A's response is
strictly greater than person B's response.

The empirical distributions come from @Burnett2014. There are four
parameters per disease: IHD, lung cancer, COPD, stroke, and LRI. In
addition, for stroke and IHD, there is a set of four parameters for each
age group from 25 to 95 in five-year increments. In addition to our
assumption that there is perfect correlation between individuals for
diseases, we assume perfect correlation between ages for diseases. I.e.,
our four quantiles per disease will be applied to all age groups.

# Results

## Health burdens in scenarios

![YLLs in the scenario minus YLLs in the baseline for
Accra.](accra_walk_yll.pdf){#results width="80%"}

## Value of information

We calculate value of information following @Jackson2017. The "value of
information" we implement is the expected value of perfect partial
information (EVPPI). It is the expected amount by which variation in the
outcome diminishes should a parameter, or set of parameters, be known
perfectly. With reference to information theory, it is the mutual
information between the parameter, or parameter set, and the outcome,
where the outcome is described only by its second moment (its variance).

The objective of learning the value of information formulated in this
way is to identify the parameter or parameters that most drive
variability (variance) in the outcome. The results guide future
research, identifying which area or areas offer the greatest benefit in
terms of prediction accuracy given further research. For example, for
our model of Accra, learning better the dose--response relationship
between physical activity and stroke (incidence) would likely increase
precision of our estimate of total years of life lost (YLL) more so than
learning any other parameter (Figure [8](#evppi){reference-type="ref"
reference="evppi"}).

![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
"all cause" and "neoplasm".](accra_walk_evppi.pdf){#evppi width="70%"}

::: {.appendix}
# ITHIM with parameters

![ITHIM-R workflow, with variable parameters in purple. Fixed inputs to
the model have a bold, black outline, which is solid for inputs we
consider "global" and dashed for inputs we consider "local". There are
four global inputs, which will be embedded in ITHIM-R. There are seven
local inputs, which users should provide. The graph depicts
parent--child relationships, where a child (at the head of an arrow)
depends upon all its parents (at the source(s) of the arrow(s)). To aid
visualisation, an arrow connected to the outside of a grey box indicates
that the arrow connects to each item within the box. *AT: active travel.
MET: metabolic equivalent task. pp: per person. AP: air pollution. PA:
physical activity. DR: dose--response relationship. GBD: global burden
of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of
it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter.
$\xi$ is 4D.](ithim_with_parameters.pdf){#ithim_with_parameters
width="\\textwidth"}

# Dose--response relationships {#doseresponse-relationships}

![1000 samples from PA dose--response
curves](PA_dose_response.pdf){#PADR width="90%"}

  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"}   ![100 samples from AP dose--response curves](cvd_strokeDR-sim.pdf){#APDR width="40%"}
  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------

  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](neo_lungDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](lriDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](resp_copdDR-sim.pdf){#APDR width="30%"}
  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------

::: {.landscape}
# Tabulated ITHIM equations {#tableequations}

::: {.center}
::: {.ThreePartTable}
::: {.TableNotes}
CDF: cumulative distribution function. We use generative distributions
to generate random numbers so that samples are correlated. For
$t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$
with CDF CDF$_F$. (I'd like to find an improved notation for this.)
:::

::: {#calculationstab}
  Name                                                                                                                                                                                    Description                                                                                      Label
  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------ -------------------------------------------
  *Scenario generation*                                                                                                                                                                                                                                                                    
  \...                                                                                                                                                                                                                                                                                     
  *Travel calculations*                                                                                                                                                                                                                                                                    
  $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk}\}\, \forall\, T_j:m(j)=\text{bus}$                                                                                           New walk trips to bus                                                                            walk_to_bus_and_combine_scen
  $\bar{T}_{a,g,m,s}=\sum_{j:a(j)=a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                                                     (Weighted) total travel per age, gender, mode and scenario                                       
  $\check{T}_{a,g,m,s}=\frac{\bar{T}_{a,g,m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}$                                                                                                             Relative travel per mode and scenario by age and gender                                          inj_distances\[\[1\]\]
  $\hat{T}_{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                                             Total travel per mode and scenario                                                               dist
  $\tilde{T}_{m,s}=\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                                                   Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1   inj_distances\[\[2\]\]
  *Injury calculations*                                                                                                                                                                                                                                                                    
  $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                                                         Casualty exponent                                                                                CAS_EXPONENT
  $E_{m_{\text{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                                                          Striker exponent                                                                                 STR_EXPONENT
  $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\cdot {I}_{a,g,m_{\text{vic}},m_{\text{str}}},\bar{T}_{a,g,m,s},E_{m_{\text{cas}}},E_{m_{\text{str}}}\right)$   Fatalities in scenarios by strike mode                                                           injuries
  $\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}$                                                                               Fatalities in scenarios                                                                          victim_deaths
  $R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}}$                                                                                     Road YLL-to-death ratio                                                                          GBD_INJ_YLL
  $\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}$                                                                                                      YLLs scaled from fatalities                                                                      deaths_yll_injuries
  *Pollution calculations*                                                                                                                                                                                                                                                                 
  $\dot{P}_{m}=\left\{\begin{array}{lr}\gamma_m & P=1 \\ \tilde{\gamma}_m, \tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)) & P<1 \end{array}\right.$                                      Emissions by mode                                                                                trans_emissions
  $\tilde{P}_{m,s}=\dot{P}_m\hat{T}_{m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                                                  Emissions by mode in scenario relative to baseline                                               trans_emissions
  $\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}$                                                                                                                                                  Emissions scalar for scenarios                                                                   baseline_sum
  $\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$                                                                                                                                              Background PM2.5 concentration in scenarios                                                      scenario_pm
  $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                                            On-road PM2.5 exposure ratio                                                                     on_road_off_road_ratio
  $\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                                       On-road PM2.5 exposure ratio for open vehicle                                                    on_road_off_road_ratio
  $\tilde{K}_{m\in\{\text{sub.}\},s}=C_6$                                                                                                                                                 Subway PM2.5 exposure ratio                                                                      subway_ratio
  $\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                In-vehicle PM2.5 exposure ratio                                                                  in_vehicle_ratio
  $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                                                          Ventilation rates                                                                                vent_rates
  $\tilde{V}_{i,m,s} = \sum_{j:i(j)=i,m(j)=m,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$                                                                                                           In-travel air inhaled                                                                            
  $\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                                                        In-travel PM2.5 inhalation rate                                                                  
  $\hat{V}_{i,s} = C_1\left(24-\sum_{j:i(j)=i,s(j)=s}T_j/Z_{m(j)}\right)$                                                                                                                 Non-travel air inhaled                                                                           
  $\check{W}_{i,s} = \bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})/24                                                                                                                            Total PM2.5 inhaled per hour                                                                     
  %/(\sum_m\tilde{V}_{i,m,s}+\hat{V}_{i,s})$                                                                                                                                                                                                                                               
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                                              AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                          AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                      AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                  AP dose--response--curve parameter                                                               DR_AP_LIST
  $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$                                                                                                                                                                                                                                 
  $\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=3}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=3}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}$                                             RR of disease given pollution                                                                    RR_AP_calculations
  *Physical activity calculations*                                                                                                                                                                                                                                                         
  $\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                                     Raw probability non-travel MMETs = 0 for demographic group $a,g$                                 raw_zero
  $\bar{Y}_{a,g}=\left\{\begin{array}{lr} \dot{Y}_{a,g} & \tilde{Y}=1 \\ \text{CDF}_{\text{Beta}(f(\tilde{Y},\dot{Y}_{a,g}))}^{-1}(\tilde{\theta}) & \tilde{Y}<1 \end{array}\right.$      Probability non-travel MMETs = 0 for demographic group $a,g$                                     zeros
  $\check{Y}_i = \left\{\begin{array}{lr} 0 & \bar{Y}_{a(i),g(i)} \\ \text{Sample}(\{Y_{i'}\}|_{a(i')=a(i),g(i')=g(i),Y_{i'}>0}) & 1-\bar{Y}_{a(i),g(i)} \end{array}\right.$              Non-travel MMETs                                                                                 SYNTHETIC_POPULATION\$ work_ltpa_marg_met
  $M_{i,s}=\sum_{j:i(j)=i,s(j)=s}(\chi T_j\cdot \lambda_{m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                                                              Total MMETs per person                                                                           mmets
  $\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$           Interpolated dose--response parameters                                                           rr, lb, ub
  $F_S(s_x) = \mathcal{N}(s_{x=2},(s_{x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                                                  Truncated normal function                                                                        
  $W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right)^{\dagger}$                                                                                   Relative risk of disease given PA                                                                RR_PA_calculations
  *Burden-of-disease calculations*                                                                                                                                                                                                                                                         
  $\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                                                     Combined PA and AP relative risks                                                                RR_PA_AP_calculations
  $\hat{W}_{a,g,h,s}= \sum_{i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                                                            Population-attributable fractions                                                                pif_temp
  $\bar{W}_{a,g,h,s}= (\hat{W}_{a,g,h,s=\text{scen1}}-\hat{W}_{a,g,h,s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                                                  Population-attributable fractions relative to scenario 1                                         pif_scen
  $\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                                                            GBD scaled to local population                                                                   DISEASE_BURDEN
  $\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}$                                                                                                                                    Scaled background disease for Neoplasms, IHD, Lung cancer                                        gbd_data_scaled
  $\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                                                         Combined health and PIF                                                                          yll_dfs, death_dfs
  $\hat{U}_{a,g,h=\text{road injury},o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                    Injury health burden                                                                             inj

  : All calculations in the ITHIM.
:::
:::
:::
:::

# Uncertainty in travel data

We model uncertainty in travel via the definition of "travel attributes"
(in the terminology of @Mohammadian2010). Specifically, we define the
propensity to travel in a day, and the distance travelled in a day,
given that travelled occurred. Each travel attribute is defined per
demographic group, per mode.

In contrast to other travel-related synthetic populations, we have four
demographic groups (rather than clusters) which consist of individuals
(rather than households). A crucial difference between our objectives
and the existing work on simulating populations in travel-demand models
(that I've found so far) is that we don't need trip-level data for our
calculations. Simulating trip-level data is computationally expensive
and high-dimensional in terms of uncertain parameters. For our
downstream computations, we need only summary data: travel per person
per mode. Beyond scenario generation, there is no need for us to
consider trips.

E.g. @Saadi2016 focuses on (a) synthetic population via matching
covariates and (b) the sequence of trips taken by each person. (a) isn't
a big thing for Accra; we're only using two covariates. And for (b), I
think we can focus on people as it's a lot simpler. For our purposes,
the things in this work that might be useful to us include matching
principles for (a), when we revisit it, and distributions (in particular
joint distributions) for travel. More useful to us would be
distributions for total travel, but I haven't found any in this
literature (e.g. zero-inflated compound Poisson distribution).

In contrast, my main questions are:

-   How to account for multi-mode trips.

-   What to do about zero travel for e.g. older cyclists. One
    possibility is to smooth over all modes and/or demographic groups.

-   In the ITHIM-R programme, are bus and truck drivers participants?

-   How should we include correlations between modes? This could also be
    aided by a smooth model across modes.

-   Whether there is a better method to get distance travelled per
    person than by resampling raw data.

## Method

### Summarise the travel survey

Our first step is to summarise the demographic information and the
travel survey. In Table [5](#population){reference-type="ref"
reference="population"} are the demographic data: the populations of the
city, and the number of people in the survey. In Table
[6](#calculations){reference-type="ref" reference="calculations"} are
the summary statistics from the travel survey which we will use to
generate synthetic travel. From the raw travel data (actually the
adjusted trip set, which is the raw travel plus motorcycle, bus and
truck trips), we calculate the probability to travel, $B_{a,g,m}$, as
the number of people who travelled by that mode divided by the number of
people in the demographic group. It is to these values that we assign
uncertainty through the variable PROPENSITY_TO_TRAVEL.

::: {.center}
::: {#population}
  Age      Gender     Surveyed   Accra population
  -------- -------- ---------- ------------------
  15--49   Male            279          5,847,716
  50--69   Male             56          1,016,476
  15--49   Female          328          6,360,535
  50--69   Female           69          1,110,037

  : Population by demographic group.
:::
:::

::: {.center}
::: {.ThreePartTable}
::: {.TableNotes}
:::

::: {#calculations}
  Age     Gender   Mode              Probability in raw trip set   Probability in baseline trip set
  ------- -------- --------------- ----------------------------- ----------------------------------
  15-49   Male     Bus                                    0.4014                             0.3862
  50-69   Male     Bus                                    0.4286                             0.3636
  15-49   Female   Bus                                    0.3506                             0.3495
  50-69   Female   Bus                                    0.4203                             0.4203
  15-49   Male     Taxi                                   0.0681                             0.0655
  50-69   Male     Taxi                                   0.0357                             0.0303
  15-49   Female   Taxi                                   0.0671                             0.0669
  50-69   Female   Taxi                                   0.1159                             0.1159
  15-49   Male     Walking                                0.6559                              0.631
  50-69   Male     Walking                                0.4643                             0.3939
  15-49   Female   Walking                                0.5427                              0.541
  50-69   Female   Walking                                0.4203                             0.4203
  15-49   Male     Private Car                            0.1254                             0.1207
  50-69   Male     Private Car                            0.1964                             0.1667
  15-49   Female   Private Car                            0.0549                             0.0547
  50-69   Female   Private Car                             0.058                              0.058
  15-49   Male     Bicycle                                0.0179                             0.0172
  50-69   Male     Bicycle                                     0                                  0
  15-49   Female   Bicycle                                 0.003                              0.003
  50-69   Female   Bicycle                                     0                                  0
  15-49   Male     Short Walking                               0                             0.3862
  50-69   Male     Short Walking                               0                             0.3636
  15-49   Female   Short Walking                               0                             0.3495
  50-69   Female   Short Walking                               0                             0.4203
  15-49   Male     Motorcycle                                  0                             0.0621
  50-69   Male     Motorcycle                                  0                             0.1515
  15-49   Female   Motorcycle                                  0                             0.0061
  50-69   Female   Motorcycle                                  0                             0.0145

  : Travel summary table.
:::
:::
:::

### Adjust for trip weight

We define a trip-mode weight per mode, $L_m$, so that a single trip by
mode $m$ in the baseline scenario represents $L_m$ trips in reality. If
in a scenario the mode of the trip is changed ($m'\rightarrow m$), it
retains the weight of its original mode $m'$. We use the trip-mode
weight to (a) sample trips for scenario generation and (b) scale the
probabilities to travel. Note that we don't use it as trip weights are
ordinarily used: to scale trip distance/duration. The result is a scaled
set of probability parameters: $$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$
($L_m$ should actually be the average trip-mode weight for all trips in
the scenario. For baseline, it will be the same. It can/should differ
for scenarios. Need to work out notation.)

### Smooth probabilities

We estimate smoothed probabilities for any mode that has zero or one
probability, using the rest of the probabilities for reference,
replacing $\hat{B}_{a,g,m}$. The reference to other modes is slightly
problematic if e.g. in the scenario all travel increases. It means a 0
in the baseline is not the same as a 0 in the scenario. This could be
fixed with more rigid coding.

### Resample probabilities

We resample the adjusted probability parameters of Table
[6](#calculations){reference-type="ref" reference="calculations"}. We
describe each parameter with a Beta distribution as follows:
$$\begin{aligned}
\label{pitilde}
\tilde{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\
\alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$
Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we
resample for each simulation. We sample via a quantile,
$\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate
$\tilde{B}_{a,g,m}$. We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum
to 200, making the distributions concentrated, and we constrain their
relative values to enforce the distribution's mean to equal
$\hat{B}_{a,g,m}$ (the right-most column in Table
[6](#calculations){reference-type="ref" reference="calculations"}).

### Inform individual travel

Each individual is assigned their own propensity to travel per mode,
$\delta_{i,m}$, which is a uniformly distributed random variable that is
fixed and not resampled. Whether or not individual $i$ travels by mode
$m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically,
individual $i$ travels by mode $m$ if
$\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$. (To mitigate the effects of
these fixed random variables, a large population is required, e.g. more
than 5000 individuals.)

![Distributions for propensities to travel. The propensity is the
probability that a member of a demographic group will travel by a
particular mode on a particular day. In orange are the raw probabilities
based on the synthetic trip set, here shown for the baseline scenario.
Beta distributions are defined via a pointiness parameter, set to 200,
such that $\alpha+\beta=200$.
](propensity_to_travel_200.pdf){#propensity_to_travel width="70%"}

## Extrapolating to one week of travel {#Extrapolation}

In order to extrapolate to one week of travel, we consider "trips" only
at the stage of scenario generation, and proceed with individual-level
statistics. Then, in our synthetic population, we have "duration
travelled per mode" columns. None of the core ITHIM-R functions will use
the "synthetic trip set".

To get duration per mode per person, we use, for each
demographic-group--mode combination, the probability to travel, and the
density of the total duration travelled given that travel occurred. To
get from one day to one week, consider two cases: one, each individual
follows the same pattern of travel every day. That is, we sample one
day's travel and apply it to seven days. Two, each individual follows
the travel pattern of their demographic group on each day. That is, we
sample seven times from the group's density to make up seven days of
travel.

In implementation, we blend individual and group-level travel through a
"repetitiveness" parameter, $\pi$. Each individual has two travel sets:
one from their simulated daily travel pattern, and one from the shared,
convolved distribution. These travel sets are weighted by $\pi$ and
$1-\pi$, respectively. $\delta_{i,m}$ defines the quantile taken for
both the individual density, which is scaled up to one week, and the
shared distribution, which is the convolution of the raw density seven
times.[^6] The densities are shown in Figure
[17](#raw_densities){reference-type="ref" reference="raw_densities"}.

![Raw distributions for duration travelled per day (left) from the
travel survey. Convolution to one week is shown on the right. We use
""walking"" mode for demographic group "Male",
"15--49".](raw_densities.pdf){#raw_densities width="80%"}

## Simulating a synthetic population {#synthpop}

1.  Choose a population size of individuals ${N}$. Label each individual
    $i=1,...,N$.

2.  Allocate each individual to a demographic group according to the
    proportions in the Accra population (Table
    [5](#population){reference-type="ref" reference="population"}). Then
    each demographic group has number $\tilde{N}_{a,g}$ and
    $\sum_{a,g}\tilde{N}_{a,g}=N$.

3.  Generate one standard uniform variable per person per mode,
    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
    $i$'s propensity to travel by mode $m$.

4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
    by mode $m$ in scenario $s$. It is the proportion of raw
    person--duration values that are zero. It is parametrised by
    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).

5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
    person--duration values for demographic group {$a,g$} and mode $m$
    in scenario $s$:
    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
    trip durations by mode $m$ in scenario $s$.

6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to
    have proportion $\tilde{B}_{a,g,m,s}$ zeros.

7.  Define each individual's travel duration as
    $D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$
    (see Figure [18](#distances_beta){reference-type="ref"
    reference="distances_beta"} for resulting distributions).

For sampling, we loop over steps 6--7, where each sample has unique
$\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain
variables is one per mode.

## Simulating a synthetic population with repetitiveness parameter & uncertain extrapolation {#synthpop2}

1.  Choose a population size of individuals ${N}$. Label each individual
    $i=1,...,N$.

2.  Allocate each individual to a demographic group according to the
    proportions in the Accra population (Table
    [5](#population){reference-type="ref" reference="population"}). Then
    each demographic group has number $\tilde{N}_{a,g}$ and
    $\sum_{a,g}\tilde{N}_{a,g}=N$.

3.  Generate one standard uniform variable per person per mode,
    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
    $i$'s propensity to travel by mode $m$.

4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
    by mode $m$ in scenario $s$. It is the proportion of raw
    person--duration values that are zero. It is parametrised by
    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).

5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
    person--duration values for demographic group {$a,g$} and mode $m$
    in scenario $s$:
    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
    trip durations by mode $m$ in scenario $s$.

6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of
    $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure
    [17](#raw_densities){reference-type="ref"
    reference="raw_densities"}, right).

7.  Define each individual's travel duration as
    $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$
    (see Figure [18](#distances_beta){reference-type="ref"
    reference="distances_beta"} for resulting distributions).

For sampling, we loop over steps 6--8, where each sample has unique
$\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of
uncertain variables is one per mode plus one.

![Samples of total distances travelled by mode and
scenario.](distances_beta.pdf){#distances_beta width="80%"}

## Results

![YLLs in the scenario minus YLLs in the baseline for Accra, with
motorcycle probability scalar 2 in
Accra.](accra_walk_yll_uncertain_travel.pdf){#results width="80%"}

![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
"all cause" and "neoplasm", with propensity to motorcycle =
2.](accra_walk_evppi_uncertain_travel.pdf){#evppi_mode_propensities_200
width="70%"}

# Glossary by letter

    $\text{Letter}$ Meaning                           $\text{Letter}$ Meaning               $\text{Letter}$ Meaning
  ----------------- ------------------------------- ----------------- ------------------- ----------------- --------------------------
           $\alpha$                                               $a$ age                               $A$ 
            $\beta$                                               $b$                                   $B$ raw travel probabilities
           $\gamma$ emission uncertainty                          $c$                                   $C$ ventilation constants
           $\delta$ propensity to travel                          $d$ demographic group                 $D$ 
         $\epsilon$ walk-to-bus time                              $e$                                   $E$ injury distance exponent
            $\zeta$ traffic PM2.5 fraction                        $f$                                   $F$ PA DR function
             $\eta$ PM2.5 concentration                           $g$ gender                            $G$ PA DR curves
           $\theta$ PA uncertainty                                $h$ disease                           $H$ AP DR curves
            $\iota$                                               $i$ individual                        $I$ injuries
                 $$                                               $j$ trip                              $J$ 
           $\kappa$ injury regression parameter                   $k$                                   $K$ AP exposure
          $\lambda$ travel MMETs                                  $l$                                   $L$ 
              $\mu$ mode distances                                $m$ mode                              $M$ MMETs per person
              $\nu$                                               $n$                                   $N$ population
              $\xi$ AP DR quantile                                $o$ outcome                           $O$ 
              $\pi$ travel repetitiveness                         $p$                                   $P$ vehicle emissions
                 $$                                               $q$                                   $Q$ person weights
             $\rho$ chronic disease scalar                        $r$                                   $R$ death-to-injury ratio
           $\sigma$ injury reporting rate                         $s$ scenario                          $S$ 
             $\tau$                                               $t$                                   $T$ trips
         $\upsilon$                                               $u$                                   $U$ burden of disease
             $\phi$ PA DR quantile                                $v$                                   $V$ AP per person
                 $$                                               $w$ AP DR parameter                   $W$ relative risks
             $\chi$ day-to-week travel scalar                     $x$ PA DR parameter                   $X$ injury model matrix
             $\psi$ casualty fraction of $\omega$                 $y$ year                              $Y$ individual MMETs
           $\omega$ sum of distance exponents                     $z$ PA dose                           $Z$ speeds
:::

[^1]: Ainsworth compendium 2011 sites.google.com/site/
    compendiumofphysicalactivities for walking and cycling

[^2]: For any person of age lower than 25, we set the relative risk to
    1.

[^3]: The four parameters refer, in numerical order, to
    $\alpha, \beta, \gamma$ and tmrel in @Burnett2014.

[^4]: This is a problem for the examples of Delhi and Bangalore, whose
    trip sets have car travellers under the age at which driving is
    permitted and who therefore should not contribute to striking
    distance.

[^5]: Note that the assymmetry in the offset term for the whw expression
    arises due to the absence of demographic information for striking
    vehicles. Were we to have that, the offset would read
    $\log(\bar{T}_{a,g,m_{\text{cas}},s})+(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})+\log(\bar{T}_{a,g,m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})$.
    For our case, the latter two terms simplify to
    $\log(\hat{T}_{m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})=E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s})$.

[^6]: Using the same random variable for both travel sets increases the
    variability of the resulting dataset. Conversely, using different
    random variables would result in more homogenous durations, and
    fewer zeros.