### 2022-11-29 results/multi_city/health_impacts/health_pathway_tbls.qmd
Add qmd to create pivot tables. Fix #123

---
title: "Health Burden (per 100k population) by pathway"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    page-layout: custom
execute:
  warning: false
  echo: false
---

```{r}
#| message: false

library(tidyverse)
library(rpivotTable)
library(htmlwidgets)

rel_path <- ""

deaths_pathway <- read_csv(paste0(rel_path, "deaths_pathway.csv"))

ylls_pathway <- read_csv(paste0(rel_path, "ylls_pathway.csv"))

plot_pivottable <- function(data, colname = "level1"){
  data |> filter(!str_detect(cause, "lb|ub")) |>   filter((!is.na(!!rlang::sym(colname)))) |>
    group_by(city, scenario, dose, !!rlang::sym(colname)) |>
    summarise(measure_100k = sum(measure_100k)) |> rpivotTable(
      rows = c("city"),
      cols = c("scenario"),
      aggregatorName = "Sum",
      vals = "measure_100k",
      rendererName = "Horizontal Stacked Bar Chart", 
      subtotals = FALSE,
      height = "100%", 
      overflow = "scroll"
    )
}

# shared_deaths_pathway <- SharedData$new(deaths_pathway |> pivot_longer(cols = -c(select(deaths_pathway, -contains("lev")) |> names())))

```

::: {.panel-tabset}

<!-- ## Reactable -->
<!-- ```{r} -->

<!-- data <- shared_deaths_pathway -->

<!-- htmltools::browsable( -->
<!--   tagList( -->
<!--     div( -->
<!--       div(tags$label("Filter Type", `for` = "cars-type-filter")), -->
<!--       tags$select( -->
<!--         id = "cars-type-filter", -->
<!--         onchange = "Reactable.setFilter('cars-filter-table', 'level1', this.value)", -->
<!--         tags$option("Level", value = "level1"), -->
<!--         lapply(data$name |> unique() |> as.character(), tags$option) -->
<!--       ) -->
<!--     ), -->

<!--     tags$hr("aria-hidden" = "true"), -->

<!--     data$data() |> filter(!str_detect(cause, "lb|ub")) |> filter(!is.na(level1)) |>  group_by(city, scenario, dose, level1) |> -->
<!--       summarise(measure_100k = sum(measure_100k)) |> rpivotTable( -->
<!--         rows = c("city"), -->
<!--         cols = c("scenario"), -->
<!--         aggregatorName = "Sum", -->
<!--         vals = "measure_100k", -->
<!--         rendererName = "Horizontal Stacked Bar Chart", -->
<!--         subtotals = FALSE, -->
<!--         height = "80%", -->
<!--         overflow = "scroll" -->
<!--       ) -->
<!--   ) -->
<!-- ) -->
<!-- ``` -->

## Deaths - Level 1
```{r}
plot_pivottable(deaths_pathway, "level1")
```

## Deaths - Level 2
```{r}
plot_pivottable(deaths_pathway, "level2")
```

## Deaths - Level 3
```{r}
plot_pivottable(deaths_pathway, "level3")
```

## YLLs - Level 1
```{r}
plot_pivottable(ylls_pathway, "level1")
```

## YLLs - Level 2
```{r}
plot_pivottable(ylls_pathway, "level2")
```

## YLLs - Level 3
```{r}
plot_pivottable(ylls_pathway, "level3")
```

:::