### 2022-11-29 results/multi_city/health_impacts/health_pathway_tbls.qmd
Change layout of results and remove unused code

---
title: "Health Burden (per 100k population) by pathway"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    page-layout: custom
execute:
  warning: false
  echo: false
---

```{r}
#| message: false

library(tidyverse)
library(rpivotTable)
library(htmlwidgets)

rel_path <- ""

deaths_pathway <- read_csv(paste0(rel_path, "deaths_pathway.csv"))

ylls_pathway <- read_csv(paste0(rel_path, "ylls_pathway.csv"))

plot_pivottable <- function(data, colname = "level1"){
  data |> filter(!str_detect(cause, "lb|ub")) |>   filter((!is.na(!!rlang::sym(colname)))) |>
    group_by(city, scenario, dose, !!rlang::sym(colname)) |>
    summarise(measure_100k = sum(measure_100k)) |> rpivotTable(
      cols = c("city"),
      rows = c("dose","scenario"),
      aggregatorName = "Sum",
      vals = "measure_100k",
      rendererName = "Stacked Bar Chart", 
      inclusions = list(scenario = list("BUS_SC")),
      exclusions= list(city = list("buenos_aires", "santiago")),
      subtotals = FALSE,
      height = "100%", 
      overflow = "scroll"
    )
}

```

::: {.panel-tabset}

## Deaths - Level 1
```{r}
plot_pivottable(deaths_pathway, "level1")
```

## Deaths - Level 2
```{r}
plot_pivottable(deaths_pathway, "level2")
```

## Deaths - Level 3
```{r}
plot_pivottable(deaths_pathway, "level3")
```

## YLLs - Level 1
```{r}
plot_pivottable(ylls_pathway, "level1")
```

## YLLs - Level 2
```{r}
plot_pivottable(ylls_pathway, "level2")
```

## YLLs - Level 3
```{r}
plot_pivottable(ylls_pathway, "level3")
```

:::