### 2023-03-17 code/synthetic_population/bogota_synth_pop.qmd
Add quarto doc reading the latest bogota travel survey from fc8bae5381440fcb96e6789d26d115305597e5f1
ref #128

---
title: "Synthetic Population for Bogota"
author: "Ali Abbas"
editor: visual
format:
  html:
    light: flatly
    dark: darkly
    toc: true
    embed-resources: true
reference-location: margin
citation-location: margin
execute:
  echo: false
  message: false
  warning: false
number-sections: true
crossref:
  chapters: true
---

## Load library
Load all the require pacakges
```{r}
library(rsample)
library(tidyverse)
library(summarytools)

# Set seed
set.seed(2023)

```

## Read latest bogoto travel survey with weights

Daniel has kindly added `participation weights` to each row in the travel dataset, which represents household weights

```{r, echo=TRUE}
bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")

view(dfSummary(bt |> filter(age != 9999)))

# rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
# rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
# 
# et <- bt %>% mutate(pw = round(participant_wt - 1))
# exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
# exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
# 
# view(dfSummary(exp1 |> filter(age != 9999)))
# 
# ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
# ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
# 
# ts <- full_join(rts, ets)
# ss <- left_join(rss, ess)
# 
# st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
# ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
# 
# view(dfSummary(st))
# view(dfSummary(ss))
# 
# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
# 
# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)

```