### 2023-03-21 code/synthetic_population/bogota_synth_pop.qmd
Improve layout of summarytools. Ref #128

---
title: "Synthetic Population for Bogota"
author: "Ali Abbas"
editor: visual
format:
  html:
    light: flatly
    dark: darkly
    toc: true
    embed-resources: true
# reference-location: margin
# citation-location: margin
execute:
  echo: false
  message: false
  warning: false
number-sections: true
crossref:
  chapters: true
---

# Load library

Load all the require packages

```{r}
library(rsample)
library(tidyverse)
library(summarytools)

# Set seed
set.seed(2023)
st_options(plain.ascii = FALSE)

```

# Read latest bogoto travel survey with weights

Daniel has kindly added `participation weights` to each row in the travel dataset, which represents household weights

```{r, results='asis'}
bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")

raw_dataset_without_weights <- bt |> filter(age != 9999)

print(dfSummary(raw_dataset_without_weights, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')

# dfSummary(bt |> filter(age != 9999), plain.ascii = FALSE, style = "grid")

# print(dfSummary(dfSummary(bt |> filter(age != 9999))), method = "render")

rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))

et <- bt %>% mutate(pw = round(participant_wt - 1))
exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)

print(dfSummary(exp1 |> filter(age != 9999),
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')

ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))

ts <- full_join(rts, ets)
ss <- left_join(rss, ess)

st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))

print(dfSummary(st,
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')

print(dfSummary(ss,
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')
# 
# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
# 
# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)

```