### 2023-03-23 code/synthetic_population/bogota_synth_pop.qmd
Add comaprison with census population - and remove distribution of sampled. Ref #128

---
title: "Travel Survey and Census population for Bogota"
author: "Ali Abbas"
editor: visual
format:
  html:
    light: flatly
    dark: darkly
    toc: true
    embed-resources: true
# reference-location: margin
# citation-location: margin
execute:
  echo: false
  message: false
  warning: false
number-sections: true
crossref:
  chapters: true
---

```{r}
library(rsample)
library(tidyverse)
library(summarytools)
library(arrow)
library(knitr)

# Set seed
set.seed(2023)
st_options(plain.ascii = FALSE)

```

# Compare Travel Survey and Census Population by age group

In this section, we compare population in the Travel Survey with Census population by sex and five year age groups.

## Distribution of `five year age groups`

The table below compares the proportion of population in Census and Travel Survey for Bogota by five year age groups.

| Age Group (5-year band) | Population (N) | Percent (%) | Travel Survey Population (N) | Percent (%) |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| 15-19                   | 562,016        | 9.8         | 3,775                        | 9.8         |
| 20-24                   | 718,131        | 12.6        | 4,750                        | 12.3        |
| 25-29                   | 724,454        | 12.7        | 4,296                        | 11.1        |
| 30-34                   | 651,833        | 11.4        | 3,831                        | 9.9         |
| 35-39                   | 599,112        | 10.5        | 3,662                        | 9.5         |
| 40-44                   | 519,983        | 9.1         | 3,278                        | 8.5         |
| 45-49                   | 470,334        | 8.2         | 3,100                        | 8           |
| 50-54                   | 465,196        | 8.1         | 3,395                        | 8.8         |
| 55-59                   | 423,099        | 7.4         | 3,253                        | 8.4         |
| 60-64                   | 336,100        | 5.9         | 3,000                        | 7.8         |
| 65-69                   | 246,821        | 4.3         | 2,307                        | 6           |

## Distribution of `sex`

The table below compares the distribution of population in Census and Travel Survey by `sex`

| Sex    | Population (N) | Percent (%) | Travel Survey Population (N) | Percent (%) |
|--------|----------------|-------------|------------------------------|-------------|
| Female | 2,996,444      | 52.4        | 17,423                       | 52.8        |
| Male   | 2,720,635      | 47.6        | 15,568                       | 47.2        |

**Takeaway**: We show the overall numbers and percentages for both of these sources. Overall, percentages of both sources are relatively similar and strengthen our confidence in the representativenesss of the household travel survey as a proxy for the overall census population.

# Compare Travel Survey distribution with and without household weights

Daniel has kindly added `weights` to each row in the Travel Survey dataset, which represents household weights. In this section we compare distribution based on `trip_mode`, `stage_mode` and `sex` distribution with and without the household weights. In all the tables below, the column `raw_trip_freq` shows distribution [**without**]{.underline} and `weighted_trip_freq` [**with**]{.underline} `weights`.

**Takeaway**: Similarly with Census, when we compare the distribution in Travel Survey with and without `weights`, they all are relatively similar across all three variables: `trip_mode`, `stage_mode` and `sex`.

```{r}

calc_freq <- function(x, y, mode = "trip_mode", round_to = 1){

  grx <- x |> filter(!is.na(across(mode))) |> group_by(across(all_of(mode))) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, round_to))

  gry <- y |> filter(!is.na(across(mode))) |> group_by(across(all_of(mode))) |> summarise(n = n()) |> mutate(weighted_trip_freq = round(n / sum(n) * 100, round_to))

  return(left_join(grx, gry, by = mode) |> dplyr::select(-all_of(starts_with("n"))))

}

bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")

raw_dataset_without_weights <- bt |> filter(age != 9999)

exp1 <- read_parquet("../../data/local/bogota/exp_trips.parquet")
# Remove undefined ages
exp1 <- exp1 |> filter(age != 9999)

```

## Distribution of `trip_mode`

Comparison of distribution of `trip_mode` with and without weights

```{r, results='asis'}

kable(calc_freq(raw_dataset_without_weights, exp1, mode = "trip_mode", round_to = 2))

```

## Distribution of `stage_mode`

Comparison of distribution of `stage_mode` with and without weights

```{r, results='asis'}

kable(calc_freq(raw_dataset_without_weights, exp1, mode = "stage_mode", round_to = 2))
```

## Distribution of `sex`

Comparison of distribution of `sex` with and without weights

```{r, results='asis'}

kable(calc_freq(raw_dataset_without_weights, exp1, mode = "sex", round_to = 2))
```

<!-- ## Summary of sampled travel survey - with trip_mode strata -->

<!-- ```{r, results='asis'} -->

<!-- st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode)) -->

<!-- kable(calc_freq(raw_dataset_without_weights, st, mode = "trip_mode", round_to = 2)) -->

<!-- ``` -->

<!-- ## Summary of sampled travel survey - with stage_mode strata -->

<!-- ```{r, results='asis'} -->

<!-- ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode)) -->

<!-- kable(calc_freq(raw_dataset_without_weights, ss, mode = "stage_mode", round_to = 2)) -->

<!-- ``` -->