### 2023-05-12 injury_tables.Rmd
Update scen names

---
title: "Injrury Outcomes - WHW matrices"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

```{r setup, include=FALSE}
library(knitr)
library(pracma)
library(summarytools)
library(dplyr)
library(purrr)
library(flextable)
library(ggplot2)
library(tidyverse)
library(writexl)

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')

smodes <- read_csv('data/global/modes/standardized_modes.csv')

# NOTE: the following scenario order is assumed: 'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc' and if there are less than 5 scenarios, the last scenarios can be missing but all of the scenarios in the middle need to exist. e.g. if you want the public transport and the walking scenario, you still need to create a car, motorcycling and bicycling scenario!

```

```{js echo=FALSE}
window.location.href='#Who_Hit_Whom_matrix_(normalized_to_annual_figures)';
```

```{r loadLibraries, echo = F, message = F}

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
# much better results. Your mileage may vary

```

```{r load_objects = "asis"}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
# Get names of cities from the io object
cities <- names(io)[!names(io) %in% 'scen_prop']

# Get number of scenarios
NSCEN <- length(names(io[[1]]$outcomes$whw)) - 1

# Get scenario names
scen_names <- names(io[[1]]$outcomes$whw)

# Remove extra bits of info - irrelevant outputs for injury related outputs
# This makes it the io object lighter
for (city in cities){
  io[[city]][["trip_scen_sets"]] <- NULL
  io[[city]][["inj_distances"]] <- NULL
  io[[city]][["synth_pop"]] <- NULL

  io[[city]]$outcomes$mmets <- io[[city]]$outcomes$pm_conc_pp <- NULL
}

round_to <- 1

# Set plot_mmets to F
plot_mmets <- F

sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

  return(qualified_scen_name)
}

```

```{r preprocessing}

collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
  genders <- unique(data$sex)
  if (ncol(data) > 3) {
    reformatted <- do.call(rbind,lapply(1:length(ages),
                                        function(x) t(sapply(genders,
                                                             function(y) 
                                                               if(ncol(data)>3) colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                        ))
    ))
    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
  }else{
    reformatted <- do.call(c,lapply(1:length(ages),
                                    function(x) sapply(genders,
                                                       function(y) 
                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                    )
    ))
    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
  }
}

scen_prop <- io$scen_prop
io <- io[-2]
# re-format some age and injury names
for (i in 1:length(io)) {
  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$deaths <- io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}
for (city in cities)
  for (type in c('hb','pathway_hb'))
    for (out in c('deaths','ylls'))
      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]]) # only keep two age categories

for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories

pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population))) # find number of people within two age categories
pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj') # find col number of injury data
ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
scen_names <- rownames(scen_prop)

```

# Who Hit Whom matrix (normalized to annual figures)

Who hit whom matrix for all scenarios for each case study

## Who hit whom matrix of each case study (by baseline + scenarios) {#whw}

```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}

all_whw <- list()

city_df <- list()

for (city in cities) {

  cat('\n')

  cat('###', stringr::str_to_title(stringr::str_replace(city, '_', ' ')), '\n')

  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
    # pref <- "ub"
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref) # if pref is lb or ub e.g.

    print(pref)

    whw_qn <- paste0("whw", pref_name)   # if upper and lower bounds are given, this takes one of whw, whw_ub, whw_lb
    nov_qn <- paste0("nov", pref_name)

    for (cs in names(io[[city]]$outcomes$whw)){ # Baseline, Scenario 1, Scenario 2, etc.
      # cs <- 'baseline'

      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){
        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if only data for the given name exists for nov but not for whw, only create data frame from nov part
          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
          td2$mode <- 'NOV'

          td2 <- td2 %>% dplyr::select(mode, names(.))

          td3 <- td2
        }

      }else{ # if whw exists
        td1 <- (io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% tibble::rownames_to_column("mode") # whw fatalities table
        td3 <- td1
        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if nov exists as well
          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()  # nov fatalities
          td2$mode <- 'NOV'

          td3 <- plyr::rbind.fill(td1, td2) # nov and whw fatalities
        }
      }
      td3 <- td3 %>% mutate(rowSum = rowSums(.[2:ncol(td3)], na.rm = T)) # add row sums
      td3 <- td3 %>% janitor::adorn_totals("row") # add column sums
      td3[, 2:ncol(td3)] <- round(td3[, 2:ncol(td3)], 2)

      qualified_scen_name <- get_qualified_scen_name(cs)

      cat('\n')

      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' WHW', '\n')

      cat('\n')

      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))] # remove column modes that do not appear in list of standardized modes

      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),] # remove rows with mode names that do not appear in list of standardized modes

      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))  # convert numbers to characters

      ft <- flextable(temp)   # convert table in correct format for printing to html file
      ft <- add_header_row(ft, values = c('Strike Mode', 'Casualty Mode'), colwidths = c(1, ncol(temp) - 1))

      cat(knit_print(ft))

      cat("\n")

      td <- td3
      names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_") # add 'city_' at beginning of column names
      td$measure <- ifelse(is.na(pref), "mean", pref)

      city_df[[cs]][[city]][[unique(td$measure)]] <- td   # save all data in one list called city_df

      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
    }

  }
  # create df with location information
  cityname <- city
  if (city == cities[[1]]){
    location <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
  } else {
    dummy_row <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
    location <- rbind(location, dummy_row)
  }

  # create dataframe with original fatality counts adjusted by weight and injury_reporting_rate

  # Baseline - whw
  injury_data_count <- as.data.frame(io[[city]]$injury_table$whw)
  injury_data_count <- injury_data_count %>% rename(value = count, str_mode = strike_mode) # rename columns

  # recalculate value based on weight and injury reporting rate
  injury_data_count$value <- injury_data_count$value / injury_data_count$weight / injury_data_count$injury_reporting_rate
  injury_data_count <- injury_data_count[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight

  # aggregate as age and gender column has been removed
  injury_data_count2 <- injury_data_count %>% group_by(cas_mode, str_mode) %>% summarise(value = sum(value))

  # Baseline - nov
  injury_data_count_nov <- as.data.frame(io[[city]]$injury_table$nov)

  if( length(injury_data_count_nov) > 0){
    injury_data_count_nov <- injury_data_count_nov %>% rename(value = count, str_mode = strike_mode) # rename columns

    # recalculate value based on weight and injury reporting rate
    injury_data_count_nov$value <- injury_data_count_nov$value / injury_data_count_nov$weight / injury_data_count_nov$injury_reporting_rate
    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight

    # aggregate as age and gender column has been removed
    injury_data_count_nov2 <- injury_data_count_nov %>% group_by(cas_mode, str_mode) %>% summarise(value = sum(value))

    city_baseline_count <- rbind(injury_data_count2, injury_data_count_nov2)
  } else {
    city_baseline_count <- injury_data_count2
  }

  city_baseline_count$city <- city
  city_baseline_count$scenario <- 'Baseline'
  city_baseline_count$country <- io[[city]]$location$country
  city_baseline_count$continent <- io[[city]]$location$continent

  if (city == cities[[1]]){
    baseline_counts <- city_baseline_count
  } else {
    baseline_counts <- rbind(baseline_counts, city_baseline_count)
  }

  baseline_counts$measure <- 'mean'

  ###### create casualty distance df
  dummy_distances <- io[[city]]$true_dist

  dummy_distances2 <- pivot_longer(dummy_distances, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_distance')

  baseline_pred <- dummy_distances2 %>% filter(scenario == 'Baseline')
  baseline_pred$scenario <- 'Baseline_predicted'

  dummy_distances3 <- rbind(dummy_distances2, baseline_pred)

  dummy_distances3$city <- city

  if (city == cities[1]){
    true_distances <- dummy_distances3
  } else {
    true_distances <- rbind(true_distances,dummy_distances3)
  }

  ####### create population df
  pop_df <- data.frame(model_population = sum(io[[city]]$demographic$population))
  pop_df$city <- city

  if (city == cities[1]){
    all_pop_df <- pop_df
  } else {
    all_pop_df <- rbind(all_pop_df, pop_df)
  }

  ####### create duration df based on true distances and speed
  speed_city <- io[[city]]$vehicle_inventory %>% dplyr::select(-c( "PM_emission_inventory",  "CO2_emission_inventory"))
  speed_city$city <- cityname

  if (city == cities[1]){
    speed_df <- speed_city
  } else {
    speed_df <- rbind(speed_df, speed_city)
  }

  # dummy_dur <- io[[city]]$dur
  # 
  # dummy_dur2 <- pivot_longer(dummy_dur, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_duration')
  # 
  # baseline_pred <- dummy_dur2 %>% filter(scenario == 'Baseline')
  # baseline_pred$scenario <- 'Baseline_predicted'
  # 
  # dummy_dur3 <- rbind(dummy_dur2, baseline_pred)
  # 
  # dummy_dur3$city <- city
  # 
  # if (city == cities[1]){
  #   duration <- dummy_dur3
  # } else {
  #   duration <- rbind(duration,dummy_dur3)
  # }
}

# calculate duration
duration <- left_join(true_distances, speed_df, by = c('city','stage_mode' ))
duration$mode_duration <- duration$mode_distance / duration$speed
duration <- duration %>% dplyr::select(-c(mode_distance, speed))

true_distances <- true_distances %>% rename(cas_mode = stage_mode)

```

```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
st <- list()
file_list <- list()
td <- NULL
whw_list <- list()
for (cs in names(all_whw)){ # loop through scenarios
  # cs <- names(all_whw)[1]

  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = c("mode", "measure")) %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))

  td[is.na(td)] <- 0

  td <- (td[!duplicated(td), ])

  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total')) # re-arrange order of rows

  td[is.na(td)] <- 0

  td <- td %>% mutate_if(is.numeric, round, 2)

  backup_td <- td

  colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))

  whw_lng <- reshape2::melt(td) # create df with 1 column containing all the fatality counts

  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column

  # tidy and rename some of the columns
  whw_lng <- cbind(whw_lng, col_split)
  names(whw_lng)[5] <- 'strike_mode'
  names(whw_lng)[6] <- 'city'
  whw_lng <- whw_lng %>% dplyr::select(-variable)
  whw_lng <- whw_lng %>% dplyr::filter(strike_mode != "rowSum")
  whw_lng <- whw_lng %>% rename(str_mode = mode, cas_mode = strike_mode)

  if (nrow(whw_lng[whw_lng$cas_mode == 'ar',]) > 0){
    whw_lng[whw_lng$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
  }

  if (nrow(whw_lng[whw_lng$str_mode == 'ar',]) > 0){
    whw_lng[whw_lng$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
  }

  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode)) # assign factors

  qualified_scen_name <- 'Public Transport'
  scen <- 'Scenario'

 qualified_scen_name <- get_qualified_scen_name(cs)

  whw_lng$scenario <- qualified_scen_name

  if (length(whw_list) == 0)
    whw_list <- whw_lng
  else
    whw_list <- rbind(whw_list, whw_lng)

  # f <- ggplot(data = whw_lng) +
  #   aes(x = str_mode, fill = cas_mode, weight = value) +
  #   geom_bar(position = "stack") +
  #   scale_fill_brewer(palette = "Dark2") +
  #   labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
  #   coord_flip() +
  #   facet_wrap(vars(city), scales = "free_y")
  # 
  # file_list[[cs]] <- f

  #ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)

  d <- paste0("results/multi_city/whw_matrices/interactive_plots/whw_", cs,".html")

  # htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
  #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")

  qual_name <- paste(qualified_scen_name, scen)
  st[[cs]] <- format(td, scientific = F)

}

# add country and continent information
whw_list <- left_join(whw_list, location, by = 'city')

# add original counts and rename Baseline to Baseline predicted
whw_list$scenario <- ifelse(whw_list$scenario == 'Baseline', 'Baseline_predicted',whw_list$scenario)

whw_list <- rbind(whw_list, baseline_counts)

# replace capital with lower letters
whw_list$str_mode <- tolower(whw_list$str_mode)

# update true distances scenarios
true_distances <- true_distances |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))

# add mode distances
whw_list <- left_join(whw_list, true_distances, by = c('city', 'cas_mode','scenario'))

# combine active travel modes
active_travel <- whw_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
active_travel <- active_travel %>% group_by(str_mode, measure, city, scenario, country, continent) %>% summarise(
                                  value = sum(value), mode_distance = sum(mode_distance), cas_mode ='active_travel')

# Remove total from whw long form, and write as csv
readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv')) # table containing all the fatality counts
```

## Who hit whom matrix for all case cities {#whw_all}

```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
### by casualty mode # This does contain NOV fatalities!

list_temp <- list()
index <- 1
scen_names <- unique(whw_list$scenario)
for (i in 1:length(scen_names)){
  sn <- scen_names[i]
  cat('###', sn, ' scenario', '\n')

  temp <- whw_list %>% filter(scenario == sn) %>% dplyr::select(cas_mode)

  cas_mode <- unique(temp$cas_mode)

  for (j in 1:length(cas_mode)){
    cm <- cas_mode[j]
    temp1 <- whw_list %>% filter(scenario == sn & cas_mode == cm) %>% spread(value = value, key = str_mode)

    temp1 <- temp1[,c(1, 2, 3, na.omit(match(smodes$exhaustive_list, colnames(temp1))), ncol(temp1))]

    temp1 <- temp1 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))

    list_temp[[index]] <- temp1

    index <- index + 1
    cat('\n')
    cat('#### Casualty mode: ', stringr::str_to_title(cm), '\n')
    #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))

    ft <- flextable(temp1)
    ft <- add_header_row(ft, values = c(' ', 'strike mode'), colwidths = c(3, ncol(temp1) - 3))
    ft <- merge_v(ft, j = 1:3, part = 'body')

    cat('\n')

    cat(knit_print(ft))

    cat('\n')

  }

  cat('\n')

}
```

### Change in deaths due to injury per billion km

```{r inj_100k = "asis", echo = F, message = F}
require(tibble)
overall_el <- list() # To save all cities results
overall_dd <- list() # To save all cities results

city_df <- list()
for (city in cities){

  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){

    el <- list() # To save city specific results
    # pref <- "ub"
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref)

    print(pref)

    whw_qn <- paste0("whw", pref_name)
    nov_qn <- paste0("nov", pref_name)

    for (cs in names(io[[city]]$outcomes$whw)){
      # cs <- 'baseline'
      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
        # td1: Number of fatalities in nov
        td1 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        # td2: Number of fatalities in whw
        td2 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        # td3: Number of fatalities: sum of nov and whw
        td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
        td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){

        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

      }
      # td4: distance travelled per mode
      td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()

      if (length(el) == 0){
        el <- td4 %>% dplyr::select(stage_mode)
        dd <- el
      }
      # td4: merge total fatalities to distance
      td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')

      var <- get_qualified_scen_name(cs)

      names(td4)[2] <- var

      td4[, 2] <- as.numeric(td4[,2])

      # add active travel mode
      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
                                                                  x = sum(active_travel[,2]),
                                                                  y = sum(active_travel[,3]))
      colnames(active_travel) <- colnames(td4)

      td4 <- rbind(td4, active_travel)

      td5 <- td4 # save copy for risk per population

      # Compute risk per bn km travelled
      td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 
                          1000000000, 4)

      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")

      el <- full_join(el, td4, by = c('stage_mode'))

      # City's population
      pop <- sum(io[[city]]$demographic$population)
      # Compute distance travelled per capita
      td6 <- td5
      td6[[paste0(var,"_risk")]] <- td4[2]
      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
      names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
      dd <- full_join(dd, td6, by = 'stage_mode')
    }

    print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))

    td <- el %>% dplyr::select(-contains('count'))
    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
    td$measure <- ifelse(is.na(pref), "mean", pref)
    dd$measure <- ifelse(is.na(pref), "mean", pref)

    city_df[["td"]][[city]][[unique(td$measure)]] <- td
    city_df[["dd"]][[city]][[unique(td$measure)]] <- dd

    if (length(city_df[["td"]][[city]]) == 3){
      overall_el[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
      overall_dd[[city]] <- data.table::rbindlist(city_df[["dd"]][[city]])
    }

    cat("\n")
  }
}
```

### Change in deaths due to injury per Billion km across cities

```{r inj_100k_all_cities = "asis"}
require(tibble)
td <- overall_el %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
td[is.na(td)] <- 0
td <- td %>% dplyr::select(stage_mode, sort(names(.)))
#readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
# colnames(td) = gsub("car_sc", "csc", colnames(td))
# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
# colnames(td) = gsub("w_sc", "wsc", colnames(td))
injury_risks_b <- td
# # Convert it into Billion by multiping the values by 10000
# injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
#readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
# colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
# colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
# colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
# colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
# colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
injury_risks_lng <- reshape2::melt(injury_risks_b)
col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
injury_risks_lng <- cbind(injury_risks_lng, col_split)
names(injury_risks_lng)[5] <- 'scenario'
names(injury_risks_lng)[6] <- 'city'
injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
# injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline_predicted'
# injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
# injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
# injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
# if (NSCEN == 4){
#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
# }
# if (NSCEN == 5){
#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
#     injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
# }
rd <- rename(injury_risks_lng, mode = stage_mode)
rd <- rd %>% filter(mode != 'Total')

# add country and continent information
rd <- left_join(rd, location, by = 'city')

# aggregate baseline counts by cas_mode
baseline_counts_agg <- baseline_counts %>% group_by(city,cas_mode, country, continent, measure, scenario) %>% summarise(value = sum(value)) %>% rename(mode = cas_mode)

# get true_distances into correct format
true_distances <- true_distances  %>% rename(mode = cas_mode)

# add distance information
rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))

# add baseline counts

# merge baseline count with distance
baseline_counts_km <- left_join(baseline_counts_agg, true_distances, by = c('mode', 'city', 'scenario'))

# add active travel 
active_travel <- baseline_counts_km %>% filter(mode == 'cycle' | mode == 'pedestrian')
active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure, ) %>% summarise(
                                   value = sum(value), mode_distance = sum(mode_distance), mode ='active_travel')

baseline_counts_km <- rbind(baseline_counts_km, active_travel)

# calculate value per distance
baseline_counts_km$value <- round((baseline_counts_km$value /(baseline_counts_km$mode_distance * 365))*1000000000,4)

rd <- rbind(rd, baseline_counts_km )

readr::write_csv(rd, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv')
```

### # of deaths due to injury per 100,000 people

```{r inj_100k = "asis"}
require(tibble)
overall_el_normalized <- list()
city_df <- list()

for (city in cities){

  print(city)

  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){

    el <- list() # To save city specific results
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref)

    print(pref)

    whw_qn <- paste0("whw", pref_name)
    nov_qn <- paste0("nov", pref_name)

    for (cs in names(io[[city]]$outcomes$whw)){
      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
        td1 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        td2 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))

      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
        td3 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){

        td3 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")

      }

      td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)

      if (length(el) == 0){
        el <- td4 %>% dplyr::select(stage_mode)
      }

      td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')

      var <- get_qualified_scen_name(cs)

      names(td4)[2] <- var

       # add active travel mode
      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
                                                                  x = sum(active_travel[,2]))
      colnames(active_travel) <- colnames(td4)

      td4 <- rbind(td4, active_travel)

      td5 <- td4

      td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)

      el <- full_join(el, td4, by = 'stage_mode')
    }

    el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

    print(kable(el, caption = city))
    td <- el
    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
    td$measure <- ifelse(is.na(pref), "mean", pref)

    city_df[["td"]][[city]][[unique(td$measure)]] <- td

    if (length(city_df[["td"]][[city]]) == 3){
      overall_el_normalized[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
    }

    cat("\n")
  }

}

td <- overall_el_normalized %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
td[is.na(td)] <- 0
td <- td %>% dplyr::select(stage_mode, sort(names(.)))
td <- rename(td, mode = stage_mode)
# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
# colnames(td) = gsub("car_sc", "csc", colnames(td))
# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
# colnames(td) = gsub("w_sc", "wsc", colnames(td))
injury_risks_per_100k <- reshape2::melt(td)
col_split <- stringr::str_split(injury_risks_per_100k$variable, "_", simplify = TRUE, n = 2)
injury_risks_per_100k <- cbind(injury_risks_per_100k, col_split)
names(injury_risks_per_100k)[5] <- 'scenario'
names(injury_risks_per_100k)[6] <- 'city'
injury_risks_per_100k <- injury_risks_per_100k %>% dplyr::select(-variable)
injury_risks_per_100k$scenario <- as.character(injury_risks_per_100k$scenario)
# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline_predicted'
# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
# injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
# if (NSCEN == 4){
#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
# }
# if (NSCEN == 5){
#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
# }

# add distance, population and location information
injury_risks_per_100k <- left_join(injury_risks_per_100k, true_distances, by = c('mode', 'city','scenario'))
injury_risks_per_100k <- left_join(injury_risks_per_100k, all_pop_df, by = c('city'))
injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')

# add counts but first divide by population
baseline_counts_pop <- left_join(baseline_counts_agg, pop_df, by = c('city'))

# add active travel 
active_travel <- baseline_counts_pop %>% filter(mode == 'cycle' | mode == 'pedestrian')
active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure ) %>% summarise(
                                   value = sum(value), model_population = sum(model_population), 
                                   mode ='active_travel')

baseline_counts_pop <- rbind(baseline_counts_pop, active_travel)

baseline_counts_pop$value <- round(baseline_counts_pop$value /baseline_counts_pop$model_population *  100000, 2)
baseline_counts_pop <- left_join(baseline_counts_pop, true_distances, by = c('city', 'mode','scenario'))

injury_risks_per_100k <- rbind(injury_risks_per_100k, baseline_counts_pop)

# add country and continent information
# injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')

readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv')
```

### True distances in a single file
```{r }
true_distances_city <- list()
for (city in cities) { # Loop for each city
  true_distances_city[[city]] <- io[[city]]$true_dist %>% 
    pivot_longer(!stage_mode, names_to = "scenario", values_to = "distance") %>% 
    mutate(city = city)
}
true_distances_df <- bind_rows(true_distances_city) # %>% 
  # mutate(scenario = case_when(
  #   scenario == "Scenario 1" ~ "Bicycling", 
  #   scenario == "Scenario 2" ~ "Car",
  #   scenario == "Scenario 3" ~ "Public Transport",
  #   scenario == "Scenario 4" ~ "Motorcycling",
  #   scenario == "Scenario 5" ~ "Walking",
  #   scenario == "Baseline" ~ "Baseline"))
readr::write_csv(true_distances_df, 'results/multi_city/whw_matrices/distances.csv')
```

### Change in deaths due to injury per 100 million hours

```{r inj_100k = "asis"}
require(tibble)
overall_el <- list() # To save all cities results
overall_dd <- list() # To save all cities results

city_df <- list()
for (city in cities){

  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){

    el <- list() # To save city specific results
    # pref <- "ub"
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref)

    print(pref)

    whw_qn <- paste0("whw", pref_name)
    nov_qn <- paste0("nov", pref_name)

    for (cs in names(io[[city]]$outcomes$whw)){ # loop through scenarios
      # cs <- 'Baseline'
      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){ # if there exist nov and whw matrices
        # td1: Number of fatalities in nov
        td1 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        # td2: Number of fatalities in whw
        td2 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        # td3: Number of fatalities: sum of nov and whw
        td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
        td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){

        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'

      }
      # td4: duration travelled per mode
      td4 <- io[[city]]$dur %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()

      if (length(el) == 0){
        el <- td4 %>% dplyr::select(stage_mode)
        dd <- el
      }
      # td4: merge total fatalities to distance
      td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')

      var <- get_qualified_scen_name(cs)

      names(td4)[2] <- var

      td4[, 2] <- as.numeric(td4[,2])

      # add active travel mode
      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
                                                                  x = sum(active_travel[,2]),
                                                                  y = sum(active_travel[,3]))
      colnames(active_travel) <- colnames(td4)

      td4 <- rbind(td4, active_travel)

      # Compute risk per 100 million hours travelled
      td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 
                          100000000, 4)

      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")

      el <- full_join(el, td4, by = c('stage_mode'))
    }

    print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))

    td <- el %>% dplyr::select(-contains('count'))
    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
    td$measure <- ifelse(is.na(pref), "mean", pref)

    city_df[["td"]][[city]][[unique(td$measure)]] <- td

    if (length(city_df[["td"]][[city]]) == 3){
      overall_el[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
         }

    cat("\n")
  }
}
```

### Change in deaths due to injury per 100 million hours across cities

```{r inj_100k_all_cities = "asis"}
#require(tibble)
td <- overall_el %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
td[is.na(td)] <- 0
td <- td %>% dplyr::select(stage_mode, sort(names(.)))

# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
# colnames(td) = gsub("car_sc", "csc", colnames(td))
# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
# colnames(td) = gsub("w_sc", "wsc", colnames(td))
injury_risks_100mil_hours <- td

# colnames(injury_risks_100mil_hours) = gsub("bi_sc", "bisc", colnames(injury_risks_100mil_hours))
# colnames(injury_risks_100mil_hours) = gsub("bus_sc", "bssc", colnames(injury_risks_100mil_hours))
# colnames(injury_risks_100mil_hours) = gsub("car_sc", "csc", colnames(injury_risks_100mil_hours))
# colnames(injury_risks_100mil_hours) = gsub("mc_sc", "mcsc", colnames(injury_risks_100mil_hours))
# colnames(injury_risks_100mil_hours) = gsub("w_sc", "wsc", colnames(injury_risks_100mil_hours))
injury_risks_lng_hours <- reshape2::melt(injury_risks_100mil_hours)
col_split <- stringr::str_split(injury_risks_lng_hours$variable, "_", simplify = TRUE, n = 2)
injury_risks_lng_hours <- cbind(injury_risks_lng_hours, col_split)
names(injury_risks_lng_hours)[5] <- 'scenario'
names(injury_risks_lng_hours)[6] <- 'city'
injury_risks_lng_hours <- injury_risks_lng_hours %>% dplyr::select(-variable)
cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
injury_risks_lng_hours$scenario <- as.character(injury_risks_lng_hours$scenario)
# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline_predicted'
# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
# if (NSCEN == 4){
#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
# }
# if (NSCEN == 5){
#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
#     injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'wsc',]$scenario <- 'Walking'
# }
rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
rd_hours <- rd_hours %>% filter(mode != 'Total')

# add country and continent information
rd_hours <- left_join(rd_hours, location, by = 'city')

# add distance information
rd_hours <- left_join(rd_hours, true_distances, by = c('city','mode','scenario'))

# get true_distances into correct format
duration <- duration  %>% rename(mode = stage_mode)
# duration[duration$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
# duration[duration$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
# duration[duration$scenario == 'Scenario 2',]$scenario <- 'Car'
# duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
# duration[duration$scenario == 'Scenario 5',]$scenario <- 'Walking'

# add duration information
rd_hours <- left_join(rd_hours, duration, by = c('mode', 'city','scenario'))

# add counts but first divide by duration
baseline_counts_dur <- left_join(baseline_counts_agg, duration, by = c('mode', 'city', 'scenario'))

# add active travel 
active_travel <- baseline_counts_dur %>% filter(mode == 'cycle' | mode == 'pedestrian')
active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure ) %>% summarise(
                                   value = sum(value), mode_duration = sum(mode_duration), 
                                   mode ='active_travel')

baseline_counts_dur <- rbind(baseline_counts_dur, active_travel)

baseline_counts_dur$value <- round((baseline_counts_dur$value /(baseline_counts_dur$mode_duration * 365))*100000000,4)
baseline_counts_dur <- left_join(baseline_counts_dur,true_distances, by = c('city', 'mode','scenario'))

rd_hours <- rbind(rd_hours, baseline_counts_dur )

readr::write_csv(rd_hours, 'results/multi_city/whw_matrices/injury_risks_per_100million_h_lng.csv')
```

### Write all injury tables in a single excel sheet
```{r}
# Export injury datasets in a single excel sheet
path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
 writexl::write_xlsx(list(whw = whw_list %>% filter(str_mode != 'Total'),
                          injury_risks_per_billion_kms = rd,
                          injury_risks_per_100k  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
                          injury_risks_per_100million_h = rd_hours,
                          true_distances = true_distances_df),
                     path = path_inj_wb)
```