### 2023-08-23 summary_tables.Rmd
updated use of reference scenario and added comments

---
title: "Summary Tables"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}

# Script summarises mode splits, distances and durations travelled and looks at health outcomes for the different air pollution, injury and physical activity pathways but also by disease

# WARNING: This script only works if ITHIM-Global has been run in constant mode!

#### It produces the following output documents:

### - html file containing the following information:

# ( - table displaying highest propensity (%) for each distance category, for a given mode. - CURRENTLY commented out as not relevant for BOGOTA, GLOBAL, LATAM or AFRICA_INDIA scenario definitions)

## Mode split
# - for each city a table is displayed showing the Baseline mode split for each distance category

# - for each city a table is given showing the mode split for each scenario

# - one table showing the baseline mode split for all cities

## Distances travelled
# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey by mode for all scenarios (incl Baseline)

# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities

# - city specific distance tables showing total yearly distances in km travelled by the total  population considered in the model for each mode for all scenarios (incl Baseline)

# - one table giving the city specific total yearly baseline distances in km travelled by the total population for each mode for all cities

# - city specific distance tables showing total daily distances in km travelled by the total population for each mode for all scenarios (incl Baseline)

# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities

## Travel duration
# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey by mode for all scenarios (incl Baseline)

# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective travel surveys for all cities

## Health outcomes - injury, air pollution and physical activity
# - one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category

# - one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group

## Health outcomes - disease
# - one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease

### - output .csv/.xlsx files containing the following information:

# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people, for each city, disease, scenario, age and sex category. It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and, the age and age and sex specific population levels.

# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv)

# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv)

# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv)

#### The script performs the following steps:

# - set up helper function to re-format large numbers and re-name scenarios

# ???????????????

# call get_scen_settings.R function to find the Baseline mode splits for each distance category for each city

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F, message = F)

```

```{r loadLibraries, echo=F, message=FALSE}
suppressWarnings({

library(summarytools)
library(knitr)
library(summarytools)
library(tidyverse)
library(gt)
library(pracma)
library(data.table)
library(writexl)
})

options(dplyr.summarise.inform = FALSE)

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
                                            # much better results. Your mileage may vary

```

```{r helper_functions, echo = F, message = F}
# Function to format numbers with million (M) and billion (B) suffixes
# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
comprss <- function(tx) { 
  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
        c("","K","M","B","T")[div] )
}

# function to rename scenarios
get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'base' | cs == 'baseline' | cs == 'Baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

  return(qualified_scen_name)
}

# define function to sum the columns of the data, print the data to html and add a caption
sum_and_round_and_print <- function(data, text = '') {
  data <- lapply(data, function(x) round(x, round_to))
  data <- lapply(data, function(x) rbind(x, Total = colSums(x)))
  for (city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

# define function to round the data and print it to html with a caption
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

```

```{r load_objects = "asis", echo = F, message = F}

# read in io object which is produced at the end of an ITHIM-Global run
# Assumes that multi_city_script.R has been run 
io <- readRDS("results/multi_city/io.rds")

# Get names of cities from the io object
cities <<- names(io)[!names(io) %in% 'scen_prop']
NSCEN <- nrow(io$scen_prop)

# find reference scenario
ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
reference_scenario <- get_qualified_scen_name(ref_scen) # update scen name

# Read trip_order
trip_order <- read_csv("data/global/trips/mode_order.csv")

# define which decimal place to round to
round_to <- 1

```
# **1st option**
## Scenario definition 
<!-- Table displays highest propensity for each distance category, for a given mode. -->

<!-- Table displays highest propensity (%) for each distance category, for a given mode. -->

<!-- In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown. -->

Distance categories are defined as:

0-2 km = \{trip distance < 2km\}

2-6 km = \{2km <= trip distance < 6km\}

6+ km = \{6km <= trip distance\}

```{r scen_prop = "asis"}

scen_prop <- round(io$scen_prop, round_to)

rownames(scen_prop) <- paste("sc",rownames(io$scen_prop),  sep = "_")

orig_scen_names_updated <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))

# # rename scenarios
# rownames(scen_prop) <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))

# kable(scen_prop, headings = "Scenario Proportions")

```

## City mode split for each distance category

```{r message=F, error=F, warning=FALSE}
# call get_scen_settings function to find the mode split for each distance category for each city
get_scen_settings <- ithimr::get_scenario_settings(cities)

```

## Mode split for each scenario by city
Case study specific trip proportions by mode, for baseline and scenarios

```{r load_tidyverse, echo = F, message = F}
suppressWarnings({
  require(tidyverse)  
})
```

```{r trip_mode_dist = "asis", echo = F}
# find the modal share of for all cities and scenarios

trip_prop <- list()

for (city in cities) { # Loop for each city

  # extract the trip data for all scenarios
  df <- io[[city]]$trip_scen_sets

  # find the number of unique trips in the Baseline
  u_trips <- df %>% dplyr::filter(scenario == "baseline") %>% 
    summarise(uid = n_distinct(trip_id)) %>% as.numeric()

  # find the proportion of trips made by each mode in each scenario
  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
    group_by(trip_mode, scenario) %>% 
    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
    spread(key = trip_mode, value = p) %>% 
    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))

  # find and update scenario names
  scen_names <- td$scenario
  scen_names_updated <- unlist(lapply(scen_names, FUN=get_qualified_scen_name))

  # take transpose
  td <- as.data.frame(t(td))

  # rename columns by using updated scenario names
  names(td) <- scen_names_updated

  # drop first column and re-order by pre-defined mode order
  td <- td[-1, ]
  x <- match(row.names(td), trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]

  # convert row names into column of data and create new dataframe
  td1 <- td %>% rownames_to_column()

  # re-name column
  names(td1)[1] <- 'stage_mode'

  # populate trip_prop list with stage modes
  if (length(trip_prop) == 0){
    trip_prop <- td1 %>% dplyr::select(stage_mode)
  }

  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

  names(td1)[2] <- city

  # add proportions for all cities but only keep stage modes that appear in the trip data of all cities
  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')

  # print mode share for individual cities
  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
  cat("\n")
} # End loop city

```

### Baseline mode split for all cities (%)
```{r}
# create baseline mode share table containing information for all cities

# prep data
trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
  janitor::adorn_totals('row')

# print table to html
kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes')

```

```{r unload_tidyverse, echo = F, message = F}
suppressWarnings({
  detach("package:tidyverse", character.only = T)
})

```

## Distance tables
City specific distance tables showing average distance in km travelled by each person by mode for all scenarios (incl Baseline)

```{r trip_dist = "asis"}
# find the average distance a person spends travelling on each mode per day

trip_dist <- list()

for (city in cities) { # loop through cities

  # find number of rows / people in the synthetic population
  count_people <- nrow(io[[city]]$synth_pop)

  # extract distances and divide by number of people
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))

  # update scenario names - cycle, car, bus
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)

  # match with pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # write to .html
  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
  cat("\n")

  td1 <- td

  # initialise trip_dist
  if (length(trip_dist) == 0){
    trip_dist <- td1 %>% dplyr::select(stage_mode)
  }

  # extract baseline distances
  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

  names(td1)[2] <- city

  # join baseline mode distances for all cities
  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')

  ## repeat but only for people who are travelling
  # extract number of people travelling from the trip data rather than the synthetic population
  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips ONLY)")))

}

```

### Distance table showing average baseline distances in km travelled by each person by mode for all cities

```{r}
# create table showing average baseline distances by mode for all cities
trip_dist <- trip_dist %>% janitor::adorn_totals('row')

kable(trip_dist, caption = "Avg. baseline distance per person (km) per day across all cities")

```

## Distance tables scaled by using total model age population (per year in km)

City specific distance tables for baseline and scenarios for the entire population with age range considered in the model

```{r trip_tot_dist = "asis", message=F}
# create distance tables for the yearly distances travelled by the entire population
bl_td <- list()

for (city in cities) { # loop through cities

  # find the total city level population considered in the model only
  #count_people <- round(sum(io[[city]]$demographic$population))

  # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances
  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
    mutate_if(is.numeric, list(~round((.) * 365, 3)))

  # update scenario names
  colnames(td)[2:ncol(td)] <- scen_names_updated

  # match with pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # initialise bl_td with stage_mode
  if (length(bl_td) == 0){
    bl_td <- td %>% dplyr::select(stage_mode)
  }

  # calculate row and column totals
  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

  # extract baseline distances
  td1 <- td %>% dplyr::select(stage_mode, Baseline)

  names(td1)[2] <- city

  # join baseline distances for all cities
  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')

  # format table
  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)

  # print to html
  cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city))
  print(td1)

  cat("\n")

}

```

### Baseline total distance for all cities (per year in km)

Baseline distance travelled by entire population with ages considered in the model for all cities

```{r trip_dist_all_cities = "asis", message=F}

# create baseline yearly distance table for all cities

backup <- bl_td

# add row totals
td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')

# remove all spaces/blanks
for (i in 2:ncol(td)){
  td[, i] <- comprss(td[, i])
}

print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model")))

```

## Distance by distance category
### Tables

```{r trip_dist_mode_figs = "asis"}
# for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category

for (city in cities){ # loop through cities

  # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category
  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% 
    summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1))

 # print to html
  print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city)))
}

```

## Daily population distances (per day in km) for ages considered in model
### Tables

```{r trip_dist_mode_figs = "asis"}

for (city in cities){ # loop through distances

  # extract total daily distances travelled by the entire population with ages considered in the model
  df <- io[[city]]$true_dist

  names(df)[-1] <- scen_names_updated

  print(kable(df, caption = paste("Model age population distance by mode per day for  ", city)))
}

```

## Duration tables

City specific duration tables for baseline and scenarios (per day in minutes)

```{r trip_dur = "asis"}
# create tables showing the average duration each person spends travelling per mode

trip_dur <- list()
l <- list()

for(city in cities){ # loop through cities
  count_people <- nrow(io[[city]]$synth_pop)

  # extract the duration for the synthetic population
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))

  # update scenario names
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)

  # order by pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # add output to list
  l[[city]] <- td

  td1 <- td

  # initialise list for all cities
  if (length(trip_dur) == 0){
    trip_dur <- td1 %>% dplyr::select(stage_mode)
  }

  # extract baseline data
  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

  names(td1)[2] <- city

  # create one baseline table containing all city information
  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')

  # print to html
  print(kable(td, caption = paste("Duration table (mins) per day for model synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))

  # repeat but only counting people who are actually taking a trip
  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id)) # find unique trip ids rather than rows in synthetic population
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Duration table (mins) per day based on trips taken in travel survey for ", city, "( ", count_people, " ) per person (people with trips)")))
  cat("\n")
}

```

### Baseline average duration (mins) per day for all cities

```{r}
# Average baseline duration travelled on each mode for all cities for all people in synthetic population

# calculate the row sum
trip_dur <- trip_dur %>% janitor::adorn_totals('row')

kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities")

```

## Health outcomes

### Change in YLLs compared to reference scenario {#change_YLL}

```{r scen_prop = "asis"}

cat(paste0('Reference scenario:  ', reference_scenario))

# Calculate the changes in YLL compared to the reference scenario by age groups

# find injury death column
#injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == paste0(scen_names[2],'_deaths_inj'))

# find ap death columns
# ap_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))

# find ap ylls columns
ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))

# find pa death columns
# pa_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))

# find pa yll columns
pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))

# # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario
# hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))]
# 
# # find the unique scenario names in the hb columns
# inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns
# # extract unique scenario names
# hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1]))
# # update scenario names
# hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name))

# from the known scenarios, remove the reference scenario
all_orig_scen_upd <- c('Baseline',orig_scen_names_updated)
hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario]

# calculate ylls by age category and scenario
# i.e. sum over all disease and remove sex 
yll_totals <- lapply(io[-2],function(x){
  temp <- sapply(1:NSCEN,function(y){# loop through scenarios
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex
  })
  colnames(temp) <- hb_scen_names_upd # update scenario names
  temp
})

# add the column total, round the results and print to html
sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ")
```

### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k}

YLLs per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate change in YLLs compared to the reference scenario per 100,000 people

# extract population statistics
pop_by_age <- lapply(io[-2], function(x)
  sapply(unique(x$demographic$age), 
         function(y)
           sum(subset(x$demographic, age == y)$population)
         )
  )

# calculate ylls per 100,000 people
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs
                                             length = NSCEN)*100000
        )
  )

names(yll_rates) <- cities

# round results and print to html
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k}

Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario
```{r scen_prop = "asis"}
cat(' ')
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category

# calculate the total ylls for each scenario summing across all diseases
yll_totals <- lapply(io[-2], function(x) {
  temp <- sapply(1:NSCEN,function(y){
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
  })
  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
                          function(z)paste0(z[2], '_', z[1]))
  colnames(temp) <- orig_scen_names_updated
  temp
})

# calculate the YLLs per 100,000 people
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1,
                                    function(z) paste0(z[2],'_',z[1])), 
                              rownames(yll_totals[[x]])),] /
          t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000,
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                             length = NSCEN) * 100000))

names(yll_rates) <- cities

# round and print ylls per 100,000 people to html
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury}

```{r}

cat(paste0('Reference scenario:  ', reference_scenario))

# calculate change in YLLs due to injury

# calculate change in YLLs due to injury by summing across all age categories for the injury results
injury_totals <- lapply(io[-2], function(x) {
  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  injury_col <- grep("inj", names(xx))[1] 
  xxx <- sapply(1:NSCEN, function(y)
    sapply(sort(unique(xx$age_cat)), function(z)
      sum(xx[xx$age_cat == z, injury_col - 1 + y])))
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate the total, round the results and print to html
sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ")
```

### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories

# sum the YLLs for each scenario by age category
pa_totals <- lapply(io[-2],function(x){
  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN, function(y){
    xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]])
    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
  })
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate column sums, round the results and print to html
sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
```

### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP}
```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLS compared to the reference scenario due to changes in AP

# sum the YLLs by scenario and age category
ap_totals <- lapply(io[-2],function(x){
  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN,function(y) {
    xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]])
    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
  })
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate column sums, round the results and print to html
sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
```

### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k}

YLLs per 100,000 people by age group by scenario - injury

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury
injury_rates <- lapply(cities, function(x)
  rbind(
    injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000,
    Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                            length = NSCEN) * 100000)
  )
names(injury_rates) <- cities

# round results and print to html
round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ")
```

### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA
pa_rates <- lapply(cities, function(x)
  rbind(
    pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000,
    Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                        length = NSCEN) * 100000))
names(pa_rates) <- cities

# round results and print to html
round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ")
```

### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP
ap_rates <- lapply(cities, function(x)
  rbind(
    ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000,
    Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]),
                                          length = NSCEN) * 100000))
names(ap_rates) <- cities

# round results and print to html
round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ")
```

## BY DISEASE

Change in total deaths compared to the reference scenario (for the city based on real population size) by scenario

### Change in deaths by disease compared to the reference scenario {#change_death_disease}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in deaths compared to the reference scenario by disease

# sum the total deaths by scenario
disease_totals <- lapply(io[-2],function(x){
  # extract disease counts
  xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) 
  xxx <- sapply(1:NSCEN,function(y){ # sum over scenarios
    colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)])
  })
  colnames(xxx) <- hb_scen_names_upd

  # find scenario name of row names
  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))

  # update row names
  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_deaths_"), "", rownames(xxx))

  xxx
})

# sum, round and print the results 
sum_and_round_and_print(disease_totals,"Change in deaths by disease in ")
```

### Change in deaths by disease compared to the reference scenario per 100,000 people {#change_death_disease_100k}

Change in deaths per 100,000 people by scenario in comparison to the reference scenario

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the change in deaths compared to the reference scenario per 100,000 people

# divided total number of deaths by population
disease_rates <- lapply(cities, function(x)
  rbind(
    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), 
                                               length = NSCEN)*100000))
names(disease_rates) <- cities

# round results and print to html
round_and_print(disease_rates,"Change in deaths by disease per 100,000 people in ")
```

### Change in total YLLs by disease compared to the reference scenario {#change_yll_disease}

Change in total YLLs by disease compared to the reference scenario (for the city based on real population size, for ages considered within the ITHIM model)

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# sum the total ylls by scenario
disease_totals <- lapply(io[-2],function(x){
  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN,function(y) {
    colSums(xx[,seq(y + 2, ncol(xx), by = NSCEN)])
  })
  colnames(xxx) <- hb_scen_names_upd

  # find scenario name of row names
  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))

  # update row names
  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_ylls_"), "", rownames(xxx))
  xxx
})

# sum and round results, print to html
sum_and_round_and_print(disease_totals,"Change in YLL by disease in ")
```

### Change in YLL due to disease compared to the reference scenario per 100,000 people {#change_yll_disease_100k}

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the change in YLL by disease by dividing by the population
disease_rates <- lapply(cities, function(x)
  rbind(
    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]),
                                               length = NSCEN) * 100000))
names(disease_rates) <- cities

# round and print results to html
round_and_print(disease_rates,"Change in YLL by disease per 100,000 in ")
```

```{r, echo=FALSE, results='hide', message=FALSE}

# Export all health results

# hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results
# pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA)

# initialise variables
health <- list()
measure <- c("ylls", "deaths")
pathway <- c("hb", "pathway_hb")

for (k in measure) { # loop for ylls or deaths
  for (j in pathway) { # loop for hb or pathway_hb
    for (i in names(io[-2])) { # loop for cities

      # Temporal dataset for health outcomes
      temp <- io[[i]]$outcomes[[j]][[k]]

      # Population by age, and age-sex
      pop_by_age <- io[[i]]$demographic %>% group_by(age) %>% 
        summarize(pop_age = sum(population))
      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>% 
        summarize(pop_age_sex = sum(population))

      # get the health burden into the correct format by creating a long format and adding additional columns
      # create long format
      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure", 
                                      -sex, -age_cat) %>% 
        mutate(city = i, # create city column
               scenario = case_when( # create scenario column
                 grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'),
                 grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'),
                 grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'),
                 grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle')
               ),
               level1 = case_when( # create level 1 column
                 grepl("all_cause", cause) ~ "L1: All Cause",
                 grepl("inj", cause) ~ "L1: RTI"
               ),
               level2 = case_when( # create level 2 column
                 grepl("total_cancer", cause) ~ "L2: Cancer",
                 grepl("CVD", cause) ~ "L2: CVD",
                 grepl("respiratory", cause) ~ "L2: Respiratory",
                 grepl("T2D", cause) ~ "L2: Other",
                 grepl("total_dementia", cause) ~ "L2: Other",
                 grepl("Parkinson", cause) ~ "L2: Other",
                 grepl("inj", cause) ~ "L2: RTI"
               ),
               level3 = case_when( # create level 3 column
                 grepl("IHD", cause) ~ "L3: IHD",
                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
                 grepl("COPD", cause) ~ "L3: COPD",
                 grepl("stroke", cause) ~ "L3: stroke",
                 grepl("T2D", cause) ~ "L3: T2D",
                 grepl("LRI", cause) ~ "L3: LRI",
                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
                 grepl("total_dementia", cause) ~ "L3: total_dementia",
                 grepl("myeloma", cause) ~ "L3: myeloma",
                 grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia",
                 grepl("Parkinson", cause) ~ "L3: Parkinson",
                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
                 grepl("inj", cause) ~ "L3: RTI"
               ),
               dose = case_when( # create dose column
                 grepl("_pa_ap_", cause) ~ "PA and AP",
                 grepl("_pa_", cause) ~ "PA",
                 grepl("_ap_", cause) ~ "AP",
                 grepl("_inj", cause) ~ "RTI"
               )
        ) %>% # End mutate 
        # join with demographic information
        left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
        # create new columns showing burden per 100000 people by age or by age and sex
        mutate(metric_100k = measure / pop_age * 100000, 
               metric_100k_sex = measure / pop_age_sex * 100000)

    } # Loop for each city
  } # Loop for each pathway
} # Loop for each measure

# for each health burden and pathway join all cities into one dataframe
ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))

# # Export ylls HB
# output_file <- "results/multi_city/health_impacts/data.xlsx"
# writexl::write_xlsx(list(ylls = ylls,
#                          ylls_pathway = ylls_pathway,
#                          deaths = deaths,
#                          deaths_pathway = deaths_pathway), 
#                     path = output_file)

# Export ylls HB
write.csv(ylls,
            "results/multi_city/health_impacts/ylls.csv",
            row.names = F)
# Export ylls PATHWAY_HB
write.csv(ylls_pathway,
            "results/multi_city/health_impacts/ylls_pathway.csv",
            row.names = F)
# Export deaths HB
write.csv(deaths,
            "results/multi_city/health_impacts/deaths.csv",
            row.names = F)
# Export deaths PATHWAY_HB
write.csv(deaths_pathway,
            "results/multi_city/health_impacts/deaths_pathway.csv",
            row.names = F)

```