### 2023-08-25 summary_tables_PA_AP.Rmd

---
title: "AP/PA Tables ITHIM Global"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: true
    toc_depth: 5
    toc_float: true
  word_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')

# Script to summarise the PM2.5 personal exposure levels, PM2.5 and CO2 concentrations and physical activity levels

# This script looks at individual level PM2.5 exposure for all scenarios, it summarise the PM2.5 and CO2 concentration levels for all scenarios and summarises the individual physical activity levels for all scenarios.

#### It produces the following output documents:

### - html file containing the following information:

# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario

# - one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario

# - one table detailing the percentage Baseline PM2.5 emissions attributed to each transport mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

# - one table detailing the Baseline CO2 concentration levels for each scenario for all cities

# - boxplots for each city of individual mMET physical activity levels for each scenario

# - one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario

# - one table detailing the average mMETs levels for each scenario for all cities

# 2nd Option:

# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario

# - one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

# - one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario. Table also contains CO2 concentrations for each scenario

# - boxplots for each city of individual mMET physical activity levels for each scenario

### - output .csv/.xlsx files containing the following information:

# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city

# - desc_stats.xlsx file (results/multi_city/AP/desc_stats.xlsx) containing the following tabs:

#   - PM_2.5_emission_inventory: table detailing the percentage Baseline PM2.5 emissions attributed to each transport mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

#   - CO2_conc: table detailing the CO2 concentration levels for each scenario for all cities

#   - summary_stats_PM2.5_CO2: summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city

#### The script performs the following steps, assuming that the ITHIM-Global model has been run in CONSTANT mode and the ithim_objects object has been saved in "results/multi_city/io.rds":

# - read in input parameter values and define function that renames the existing scenarios

# - create boxplots of individual PM2.5 levels for each scenario, one for each city, and write to the .html file

# - create one table for each city detailing the individual PM2.5 exposure levels incl. the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario

# - create a single table detailing the percentage Baseline PM2.5 emissions attributed to each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

# - create a single summary table showing the CO2 concentrations in each scenario for all cities

# - create boxplots for each city showing the individual mMET levels for each scenario

# - create one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario

# - create a single table detailing the average mMETs levels for each scenario for all cities

# - 2nd option:

# - create boxplots for each city of daily individual PM2.5 exposure levels for each scenario

# - create one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

# - create one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario. Table also contains CO2 concentrations for each scenario

# - create boxplots for each city of individual mMET physical activity levels for each scenario

```

# Introduction
These are the summary tables of the following items

1) Individual PM2.5 exposure levels and total PM2.5 concentrations for baseline and scenarios
2) Overall CO2 emissions for baseline and scenarios
3) Individual physical activity levels (mMETs) for baseline and scenarios

```{r, message=FALSE, warning=FALSE}
#library(INLA)     #loading the INLA package
library(ggplot2)  #loading ggplot package for plotting graphs
library(knitr)     
library(tidyverse)
library(readxl)
#library(plotly)
#library(htmltools)

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}

###### Read in input parameters

# Assumes that multi_city_script.R has been run  
# read in input file
io <- readRDS("results/multi_city/io.rds")

# Get names of cities from the io object
cities <- names(io)[!names(io) %in% 'scen_prop']

# function to convert scenario names
get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'base' | cs == 'Baseline' | cs == 'baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

  return(qualified_scen_name)
}

# initialise parameters using first city in list
city <- cities[1]
scen_length <- length(io[[city]]$ithim_run$scenario_names) - 1

# input parameter file name
input_parameter_file <<- io[[city]]$ithim_run$input_parameter_file

# scenario and reference scenario definitions
scenario_name <- io[[city]]$ithim_run$scenarios_used
ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
reference_scenario <- get_qualified_scen_name(ref_scen)

compute_mode <- io[[city]]$ithim_run$compute_mode

# read in local input parameter inputs
all_inputs <- read_excel(input_parameter_file, sheet = "all_city_parameter_inputs")
all_inputs[is.na(all_inputs)] <- ""
all_inputs <- as.data.frame(all_inputs)

# get input parameters into correct format
parameter_names <- all_inputs$parameter
parameter_starts <- which(parameter_names!='')
parameter_stops <- c(parameter_starts[-1] - 1, nrow(all_inputs)) 
parameter_names <- parameter_names[parameter_names!='']
parameter_list <- list()

# extract local parameter information
for(i in 1:length(parameter_names)){
  parameter_list[[parameter_names[i]]] <- list()
  parameter_index <- which(all_inputs$parameter==parameter_names[i]) 
  if(all_inputs[parameter_index,2]=='')  { 
    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
      city_index <- which(colnames(all_inputs)==x)
      val <- all_inputs[parameter_index,city_index]
      ifelse(val%in%c('T','F'),val,ifelse(is.numeric(val), as.numeric(val), as.character(val)))
    })
    names(parameter_list[[parameter_names[i]]]) <- cities
  }else if(all_inputs[parameter_index,2]=='constant'){
    if (compute_mode != 'sample'){
      indices <- 0
      parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
        city_index <- which(colnames(all_inputs)==x)
        val <- all_inputs[parameter_index+indices,city_index]
        ifelse(val=='',0,as.numeric(val))
      })
    }
    if(compute_mode=='sample'){ # if sampling from distribution, check that distribution parameters exist
      parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
        indices <- 1:2
        city_index <- which(colnames(all_inputs)==x)  
        val <- all_inputs[parameter_index+indices,city_index] 
        if (val[1] == '' & val[2]==''){  # if no distribution parameters given in input file, read in constant value instead
          indices <-0
          city_index <- which(colnames(all_inputs)==x) 
          val <- all_inputs[parameter_index+indices,city_index]} 
        val <- as.numeric(val)
      })
    }
    names(parameter_list[[parameter_names[i]]]) <- cities
  }else{
    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
      city_index <- which(colnames(all_inputs)==x)
      if(any(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')){
        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')
        thing <- as.list(as.numeric(c(all_inputs[parameter_starts[i]:parameter_stops[i],city_index])[sublist_indices]))
        names(thing) <- c(all_inputs[parameter_starts[i]:parameter_stops[i],2])[sublist_indices]
        thing
      }
    }
    )
    names(parameter_list[[parameter_names[i]]]) <- cities
  }
}

list2env(parameter_list, environment())

# read in global parameters
all_global_inputs <- read_excel(input_parameter_file, sheet = "all_global_parameter_inputs")
all_global_inputs[is.na(all_global_inputs)] <- ""
all_global_inputs <- as.data.frame(all_global_inputs)

# get input parameters into correct format
global_parameter_names <- all_global_inputs$parameter
global_parameter_starts <- which(global_parameter_names!='')
global_parameter_stops <- c(global_parameter_starts[-1] - 1, nrow(all_global_inputs)) 
global_parameter_names <- global_parameter_names[global_parameter_names!='']
global_parameter_list <- list()

# extract global parameter information
for(i in 1:length(global_parameter_names)){
  global_parameter_list[[global_parameter_names[i]]] <- list()
  global_parameter_index <- which(all_global_inputs$parameter==global_parameter_names[i]) 
  if(all_global_inputs[global_parameter_index,2]=='')  {

    global_parameter_list[[global_parameter_names[i]]] <- all_global_inputs[global_parameter_index,'global']

  }else if(all_global_inputs[global_parameter_index,2]=='constant'){
    if (compute_mode != 'sample'){
      global_parameter_list[[global_parameter_names[i]]] <- ifelse(all_global_inputs[global_parameter_index,'global']=='',
                                                                   0,as.numeric(all_global_inputs[global_parameter_index,'global']))
    }
    else if(compute_mode=='sample'){ # if sampling from distribution, check that distribution parameters exist
      indices <- 1:2
      val <- all_global_inputs[global_parameter_index+indices,'global'] 
      if (val[1] == '' & val[2]==''){  # if no distribution parameters given in input file, read in constant value instead
        val <- all_global_inputs[global_parameter_index,'global']} 
      val <- as.numeric(val)
      global_parameter_list[[global_parameter_names[i]]] <- val
    }
  }
}

list2env(global_parameter_list, environment())

# get parameters into correct format
dist_cat <- unlist(strsplit(gsub(" ", "", dist_cat, fixed = TRUE), "\\,"))

outcome_age_min <- as.numeric(unlist(strsplit(gsub(" ", "", outcome_age_min, fixed = TRUE), "\\,")))
outcome_age_max <- as.numeric(unlist(strsplit(gsub(" ", "", outcome_age_max, fixed = TRUE), "\\,")))
outcome_age_groups <- unlist(strsplit(gsub(" ", "", outcome_age_groups, fixed = TRUE), "\\,"))

min_age <- as.numeric(min_age)
max_age <- as.numeric(max_age)

day_to_week_scalar <- as.numeric(day_to_week_scalar)

```

# Boxplots of individual daily PM2.5 exposure levels

```{r}

#### for each city create a boxplot showing the individual daily PM2.5 exposure levels for all scenarios

# loop through cities
for (x in 1:length(cities)) {

  n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)

  # extract original scenario names and convert to updated scenario names
  col_names <- names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col]
  orig_scen_names <- sub("pm_conc_", "", col_names)
  new_scen_names  <- unlist(lapply(orig_scen_names, FUN=get_qualified_scen_name))

  # rename columns
  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names

  # get PM2.5 individual exposure level data into correct format
  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create boxplot
  y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
    geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
    labs(y = "Daily PM2.5 exposure levels", x = "Scenarios") #+

  # print to html
  print(y)

}
```

# Summary statistics of individual PM2.5 exposure levels

```{r, message=FALSE, warning=FALSE, echo=FALSE}

cat(paste0('The change in PM2.5 levels is given as the difference to the reference scenario: ', reference_scenario))

# for each city create a summary table of individual PM2.5 exposure levels giving the mean and the 5th, 50th and 95th percentile levels. Tables also show the PM2.5 concentration levels in the city for each scenario and the changes compared to the reference scenario

data_long <- NA
summary <- NA
for (x in 1:length(cities)) {

  n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)

  # rename columns with updated scenario names
  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names

  # get data into correct format
  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create summary stats
  summary <- as.data.frame(data_long %>% group_by(scenario) %>%
                             summarise('mean' = mean(pm_conc),
                                       '5th' = quantile(pm_conc, 0.05),
                                       '50th' = quantile(pm_conc, 0.5),
                                       '95th' = quantile(pm_conc, 0.9)))
  summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))

  # add scenario specific PM2.5 concentration levels 
  summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)

  # find location of reference scenario in vector of all scenarios
  loc_ref <- which(new_scen_names == reference_scenario)

  # calculate the PM2.5 changes between the reference scenario and the other scenarios
  summary$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[loc_ref], digits=2)
  names(summary)[1] <- "Scenario"

  # print to html file
  print(kable(summary, caption = cities[x]))

}
```

# Summary table of PM 2.5 emission inventories

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# Create a single table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

sl <- list()
for (x in 1:length(cities)){ # loop through cities
  print(cities[x])

  # find the modes given in the PM emission inventory for that city
  modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))

  # extract the PM emissions for each mode
  emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))

  # combine modes and emissions into one dataframe
  city_emissions <- cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
  names(city_emissions)[2] <- "emissions"

  # define modes of interest
  select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
  city_emissions$modes <- as.character(city_emissions$modes)

  # call all other modes not in list of modes of interest 'other'
  city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"

  # calculate the sum of PM emissions by mode (primarily summing across all newly classified 'other' modes)
  summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
  names(summary)[2] <- cities[x]

  # calculate the percentages of PM emissions attributed to each mode
  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
  summary <- as.data.frame(summary)

  # rename mode columns including the word percentage
  summary$modes <- paste0(summary$modes, ' (percentage)')

  # add transport share 
  summary[nrow(summary) + 1,1] <- "Transport share (percentage)"
  summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))

  # add city wide PM2.5 concentration levels
  summary[nrow(summary) + 1,1] <- "PM2.5 Conc"
  summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])

  # add to io list
  io[[cities[x]]]$summary_emission <- summary

  # create one dataframe containing all cities
  if (length(sl) == 0){
    sl <- summary
  }else{
    sl <- left_join(sl , summary)
  }

}

# print table to html
print(kable(sl))
```

# Summary table of CO2 concentrations related to transport in each scenario

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# create summary table of CO2 concentrations in each scenario
co2_conc <- list()
td <- NA

for (city in cities) { # loop through cities

  # extract CO2 concentrations
  td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% 
    as.data.frame() %>% tibble::rownames_to_column()
  names(td) <- c('Scenario', city)

  # create one dataframe containing all cities
  if (length(co2_conc) == 0)
    co2_conc <- td
  else
    co2_conc <- left_join(co2_conc, td)

}

co2_conc$Scenario <- new_scen_names

# print to html
print(kable(co2_conc))

```

# Boxplots of individual physical activity (MMETs) levels

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# create boxplots for each city showing the individual mMET levels for each scenario

limit = 100

for (x in 1:length(cities)) { # loop through cities

  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # define columns of interests
  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names

  # extract data of interest
  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create the boxplots
  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) +
    geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
    labs(y = "Weekly mMET", x = "Scenarios")+  ylim(0, limit)

  # print to html
  print(y)

}
```

# Summary statistics of individual physical activity levels (MMETs)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# create one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario

for (x in 1:length(cities)) { # loop through the cities

  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find the number of columns
  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names # assign new column names

  # create dataframe with individual mMET values
  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create summary stats
  summary <- as.data.frame(data_long %>% group_by(scenario) %>% 
                             summarise('mean' = mean(mmet),
                                       '5th' = quantile(mmet, 0.05),
                                       '50th' = quantile(mmet, 0.5),
                                       '95th' = quantile(mmet, 0.9)))
  summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))
  names(summary)[1] <- "Scenario"

  # print to html
  print(kable(summary, caption = cities[x]))

}
```

# Average baseline physical activity levels (MMETs) in each city

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# create a single table detailing the average mMETs levels for each scenario for all cities

cl <- list()
for (city in cities){ # loop through cities

  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns of interest

  # extract the mean mMET values and create dataframe
  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - scen_length):n_col], na.rm = T), 1) %>%
    as.data.frame() %>% tibble::rownames_to_column()
  names(td) <- c('Scenario', city)

  # create one dataframe containing information about all cities
  if (length(cl) == 0)
    cl <- td
  else
    cl <- left_join(cl, td)

}

cl$Scenario <- new_scen_names

# print to html 
print(kable(cl))

```

# **2nd option**

# PM2.5 and CO2 results

```{r message=FALSE, warning=FALSE, echo=FALSE}

# create PM2.5 and CO2 boxplots and tables:

# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario

# - one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities

# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario. Table also contains CO2 concentrations for each scenario

l <- list()
# Have to keep the first chunk of code to run this part

for (x in 1:length(cities)) { # loop through cities
  # print(cities[x])
  cat("\n \n")
  cat( paste0('## **', cities[x], '** \n' ))

  ## Emission inventory

  # find the mode names in the PM2.5 emissions inventory
  modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))

  # extract the mode emissions
  emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))

  # create one table with modes and corresponding PM2.5 emissions
  city_emissions <- cbind(as.data.frame(modes),
                          as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))

  names(city_emissions)[2] <- "emissions"

  # define list with modes of interest
  select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
  city_emissions$modes <- as.character(city_emissions$modes)

  # define all other modes not in list of interest as 'other'
  city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"

  # sum the emissions (mainly to find the total emissions for all newly defined 'other' modes)
  summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
  names(summary)[2] <- cities[x]

  # calculate the percentage contribution of each mode
  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
  summary <- as.data.frame(summary)

  # rename mode columns including the word percentage
  summary$modes <- paste0(summary$modes, ' (percentage)')

  # add the percentage attributed to transport
  summary[nrow(summary)+1,1] <- "Transport share (percentage)"
  summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))

  # add the total PM2.5 concentration levels
  summary[nrow(summary)+1,1] <- "PM2.5 Conc"
  summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])

  # add to io object
  io[[cities[x]]]$summary_emission <- summary

  ## PM2.5 table and plot

  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns

  # assign new scenario names to columns
  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names

  # extract data with PM2.5 exposure levels for each scenario
  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create summary statistics
  summary2 <- as.data.frame(data_long %>% group_by(scenario) %>%
                             summarise('mean' = mean(pm_conc),
                                       '5th' = quantile(pm_conc, 0.05),
                                       '50th' = quantile(pm_conc, 0.5),
                                       '95th' = quantile(pm_conc, 0.9)))
  summary2 <- cbind(summary2$scenario, round(summary2[,2:5], digits = 1))

  # add total PM concentration levels for each scenario
  summary2$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, 2)

  # calculate difference to reference scenario
  summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[loc_ref], digits = 2)
  names(summary2)[1] <- "Scenario"

  # Plot individual level PM2.5 exposure levels by scenario
  y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
    geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
    labs(y = "Daily PM2.5 Concentration", x = "Scenarios")

  # CO2
  # extract CO2 transport emissions for each scenario
  td <- round(colSums(io[[cities[x]]]$outcomes$co2_conc, na.rm = T), 1) %>%
    as.data.frame() %>% tibble::rownames_to_column()
  names(td) <- c('Scenario', 'CO2')
  td$Scenario <- new_scen_names

  # print various boxplots and tables to html
  cat("\n \n")
  cat( paste('#### Boxplots of individual PM2.5 exposure levels \n' ))
  print(y)

  cat("\n \n")
  cat( paste('#### PM2.5 emissions inventory \n' )) 
  print(kable(summary))

  cat("\n \n")
  cat( paste('#### Summary statistics of individual level PM2.5 exposure levels and total transport related CO2 emissions\n' ))
  cat(paste0('The change in PM2.5 levels is given as the difference to the reference scenario: ', reference_scenario))
  l[[cities[x]]] <- left_join(summary2, td, by = "Scenario")
  l[[cities[x]]]$city <- cities[x]
  print(kable(cbind(summary2, td), caption = cities[x]))

}

# create output Excel file containing the basic PM2.5 and CO2 exposure / concentration information
output_file <- "results/multi_city/AP/desc_stats.xlsx"
writexl::write_xlsx(list(PM_2.5_emission_inventory = sl,
                         CO2_conc = co2_conc,
                         summary_stats_PM2.5_CO2 = bind_rows(l)), path = output_file)

# create csv file (summary_stats_Pm2.5_CO2 tab in Excel file above)
write_csv(bind_rows(l), "results/multi_city/AP/desc_stats.csv")

```

# Boxplots of individual physical activity levels (MMETs)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# create boxplots for each city of individual mMET physical activity levels for each scenario

limit = 100
for (x in 1:length(cities)) { # loop through cities

  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns

  # assign new scenario names
  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names

  # extract mMETs data
  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)

  # create boxplots
  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
    labs(y = "Weekly MMET", x = "Scenarios") + ylim(0, limit)

  # print to html
  print(y)

}
```