### 2023-12-12 results_app/results_presentation.qmd

---
title: "ITHIM Results"
author: "Ali Abbas"
format: revealjs
editor: visual
---

## ITHIM Overview

### Structure

#### Figure

#### Pathways

## Outcomes

## Scenario Definition - only 5% change in trips

## Distance and Duration tables

## PA outcomes - average MMETs per peron per week

## AP Outcomes - PM2.5 exposure levels

## Health Outcomes

### YLLs and Deaths by total, sex and gender stratification - with and without interaction for three levels and covering all three paths

### Touch on unexpected results when median values of MMETs per week compared in bus and cycle scenarios

## Injury risks by population, distance and duration - overall and not stratified