### 2023-12-13 results_app/results_presentation.qmd
Add presentation on results

---
title: "ITHIM Results"
author: "Ali Abbas"
date: today
date-format: long
title-slide-attributes:
  data-background-image: "images/ithim-accra.png"
  data-background-size: contain
  data-background-opacity: "0.2"
execute:
    echo: false
    warning: false
    message: false
    cache: false
format:
  revealjs:
    standalone: false
    embed-resources: true
    width: 1600
    height: 900
    theme: quartomonothemer.scss
    slide-number: true
    fig-cap-location: top
    pdf-separate-fragments: false
    transition: slide
    background-transition: fade

output: 
  revealjs::revealjs_presentation: default

reveal_options:
  minScale: 1.0
  maxScale: 1.0
---

```{r}
library(tidyverse)
library(kableExtra)
library(modelsummary)
library(fixest)
library(gt)
library(gtExtras)
library(gghighlight)
library(quartomonothemer)
library(showtext)
```

```{r}
font_title <- "Josefin Sans"
font_text <- "Montserrat"
font_sans <- "Noto Sans" 
color_base <- "#009F8C"
color_base_light <- "#95DFD6"
color_accent <- "#B75C9D"
color_accent_light <- "#DBA6CC"
gray <- "#bebebe"
darkgray <- "#6d6d6d"

font_add_google(font_title)
font_add_google(font_text)
showtext_auto()

style_mono_quarto(
  font_title = font_title,
  font_text = font_text,
  font_sans = font_sans,
  color_base = color_base,
  color_accent = color_accent,
  color_link = color_accent,
  color_code = color_base,
  size_base = 30,
  path_scss = here::here("quartomonothemer.scss")
)
```

```{r}
# function to convert scenario names
get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'base' | cs == 'Baseline' | cs == 'baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

  return(qualified_scen_name)
}

```

```{r}

library(tidyverse)
library(plotly)

library(here)
library(tidyverse)
library(gghighlight)
library(gt)
library(gtExtras)

repo_sha <-  as.character(readLines(file.path("repo_sha")))
io <- readRDS(paste0("../results/multi_city/io_", repo_sha, "_test_run.rds"))

PIF <- read_csv("data/PIF.csv")

PIF <- PIF |> rename(Scenario = name) |> mutate(Scenario = case_when(
                 Scenario  == "sc_cycle_PIF" ~ "Bicycling",
                 Scenario  == "sc_car_PIF" ~ "Car",
                 Scenario  == "sc_bus_PIF" ~ "Public Transport",
               )
)

prop <- io$bogota$trip_scen_sets |> distinct(scenario, trip_id, .keep_all = T) |> count(scenario, trip_mode) |> mutate(freq = round(prop.table(n) * 100, 1), .by = scenario) |> mutate(pd = round(freq - freq[scenario == 'baseline'], 1)) |> filter(pd != 0)

prop[prop$scenario == "baseline",]$scenario <- "Baseline"
prop[prop$scenario == "sc_bus",]$scenario <- "Public Transport"
prop[prop$scenario == "sc_car",]$scenario <- "Car"
prop[prop$scenario == "sc_cycle",]$scenario <- "Bicycling"

scen_colours <- c("Baseline" = '#ffffbf',
                  "Bicycling" = '#abdda4',
                  "Car" = '#d7191c',
                  "Public Transport" = '#2b83ba',
                  "Motorcycle" = '#fdae61')

```

# Scenario Definition - (5% change)

## Scenario Proportions

![](images/trip-prop.png)

## Cycling Scenario

![](images/cycle-scenario.png)

## Bus Scenario

![](images/bus-transport-scenario.png)

## Car Scenario

![](images/car-scenario.png)

## Comparison

```{r}

plotly::ggplotly(ggplot(prop) +
     aes(x = trip_mode, fill = scenario, weight = pd) +
     geom_bar() +
     scale_fill_hue(direction = 1) +
    scale_x_discrete(guide = guide_axis(angle = 90)) + 
    geom_text(aes(x = trip_mode, y = round(pd, 1), label = round(pd, 1)), size = 4, 
                                 position = position_dodge()) + scale_fill_manual(values = scen_colours) +
     facet_wrap(vars(scenario)) + 
     labs(title = "Scenario definition (by mode)", y = "Percentage (%)", x = "Trip Mode") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))

```

# Distance tables - (km)

## Distance table

![](images/distance-table.png)

## Cycling Scenario distance table

![](images/distance-table-cycle-scenario.png)

## Bus Scenario distance table

![](images/distance-table-bus-scenario.png)

## Car Scenario distance table

![](images/distance-table-car-scenario.png)

# Duration tables - (minutes)

## Duration table

![](images/duration-table.png)

## Cycling Scenario duration table

![](images/duration-table-cycle-scenario.png)

## Bus Scenario duration table

![](images/duration-table-bus-scenario.png)

## Car Scenario duration table

![](images/duration-table-car-scenario.png)

## ITHIM Overview

### Structure

#### Figure

#### Pathways

## Outcomes

## Duration table

![](images/duration-table.png)

## PM Emission Inventory

```{r fig.height=5, fig.width=8}

gtb <- io$bogota$vehicle_inventory |> 
  dplyr::select(-c(speed, CO2_emission_inventory)) |> 
  filter(PM_emission_inventory !=0) |> 
  mutate(PM_emission_inventory = round(PM_emission_inventory * 100, 1)) |> 
  gt() |> 
  cols_label(stage_mode = "Mode", PM_emission_inventory = "PM Emission Inventory") |>
  gt_plt_bar_pct(PM_emission_inventory, scaled = TRUE, labels = TRUE)

gtb
```

## AP Outcomes - PM2.5 exposure levels ($\mu$/$m^3$)

```{r mmets}

mmets <- io$bogota$outcomes$mmets #|> pivot_longer(cols = ends_with("mmet"))

change_names <- which(grepl("_mmet", names(mmets), fixed = TRUE))
  names(mmets)[change_names] <- sapply(gsub("_mmet" , "", names(mmets)[change_names]), FUN = get_qualified_scen_name)

mmets <- mmets |> pivot_longer(cols = -c(1:4)) |> rename(Scenario = name)

plotly::ggplotly(ggplot(mmets, aes(x=value, fill=Scenario)) + 
                   geom_density(alpha=.3) + 
                   scale_fill_manual(values = scen_colours) +
                   xlim(0, 120) +
                   labs(title = "Cumulative Distribution of MMETs per person per week (by scenario)",
                        x = "MMETs per person per week",
                        y = "Density")
                   )

```

## PA outcomes - Marginal METs per person per week

```{r}
mmets |> 
    group_by(Scenario) |> 
    reframe(qs = c(quantile(value, prob = 0.25), median(value), mean(value))) |>
    mutate(qs = round(qs, 1)) |> group_by(Scenario) |> summarise(ld = list(qs)) |> gt() |> gt_plt_bar_stack(column = ld, labels = c("Minimum", "Median", "Mean"), palette = c("#fee8c8", "#fdbb84", "#e34a33"))

```

## Potential Impact Fraction using different censor points

```{r}

ggplotly(ggplot(PIF) +
  aes(x = group, y = value, fill = Scenario) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = scen_colours) +
  facet_wrap(vars(Scenario)) +
  labs("Potential Impact Fraction using all-cause DRF",
  subtitle = "We have used three different DRFs where censor point varies. PIF_75p where censor point is at 75th percentile, and PIF_WHO_DRL (censor point is at 35 MMETs per person per week and PIF_none (without any censor point)",
  x = "", y = "PIF (%) (+ve benefit)"))

```

## Health Outcomes

### YLLs and Deaths by total, sex and gender stratification - with and without interaction for three levels and covering all three paths

### Touch on unexpected results when median values of MMETs per week compared in bus and cycle scenarios

## Injury risks by population, distance and duration - overall and not stratified